---
output: 
  word_document:
    reference_docx: ../report_template_v1.5.dotx
---

```{r setup, include=FALSE}

# =======================================================================================
# Blue whiting catch by rectangle.Rmd
# 
# Generate stock summary report for WGWIDE
#
# 03/09/2017 first version
# 06/08/2018 UPdated and prepared for WGWIDE 2018
# 07/09/2018 Added link to SAG database for plotting. Currently reading from downloaded data. 
#            Needs to be made through live connection. 
# 16/11/2021 Adapted for Coastal States meeting
# =======================================================================================

require("knitr")
knitr::opts_chunk$set(echo = FALSE,	message = FALSE,	warning = FALSE,	comment = "",	crop = TRUE )
knitr::opts_chunk$set(fig.width=10) 

rm(list=ls())

library(tidyverse)    # tidying packages
library(readxl)       # read excel
library(lubridate)
library(sf)
library(sp)
library(RColorBrewer)
library(pander)

# library(gisland)      # devtools::install_github("einarhjorleifsson/gisland", dependencies = TRUE)

library(maps)
# library(mapdata)
# library(mapproj)
library(lubridate)
library(viridis)

# library(animation)
# library(gganimate)

# source my utils
source("../../prf/R/my utils.r")
source("../../gisland/r/geo_inside.R")

dropboxdir <- paste(get_dropbox(), "/iAdvice", sep="")

# Load dataset
load(file=paste(dropboxdir, "/rdata/iAdvice.RData",sep=""))

# Data path
datapath <- "C:/Users/Martin/Onedrive - PFA/Documents/iWGWIDE/2021/06. Data/_catch_by_rectangle"
# datapath <- "//community.ices.dk@SSL/DavWWWRoot/ExpertGroups/WGWIDE/2019 Meeting docs/06. Data/_catch_by_rectangle"

datapath2 <- "C:/Users/Martin/Havstovan/WG on mackerel distribution - 2021 WHB update report on distribution/Data/Catch data"

onedrive <- get_onedrive()

load(file.path(onedrive, "rdata/world.df.RData"))
load(file.path(onedrive, "rdata/eez.df.RData"))
load(file.path(onedrive, "rdata/fao.df.RData"))
load(file.path(onedrive, "rdata/depth200.df.RData"))
load(file.path(onedrive, "rdata/icesrectangles.df.RData"))

load(file.path(onedrive, "rdata/eez.RData"))

load(file.path(onedrive, "rdata/icesrectangles.sf.RData"))
load(file.path(onedrive, "rdata/icesrectangles.eez.sf.RData"))
load(file.path(onedrive, "rdata/eez.sf.RData"))

icesrect <- 
  icesrectangles.df %>% 
  group_by(rect) %>% 
  distinct(rect, .keep_all = TRUE) %>% 
  mutate(
    lon = (WEST + EAST)/2,
    lat  = (SOUTH + NORTH)/2
  ) %>% 
  dplyr::select(rect, lon, lat)

# icesrect <-
#   icesrectangles.df %>%
#   distinct(rect, lat=SOUTH, lon=WEST) %>%
#   dplyr::select(rect, lon, lat)
# write.csv(icesrect, file="icesrect.csv", row.names = FALSE)

# icesrect <- read.csv(file=file.path(datapath, "icesrect.csv")) %>% mutate(across(c("lat","lon"), as.numeric))

zbyrect <- 
  icesrectangles.eez.sf %>% 
  # read.csv(file=file.path(datapath, "ZbyCountryNew.csv")) %>%
  sf::st_set_geometry(NULL) %>% 
  lowcase() %>% 
  dplyr::select(rect, zone, lon, lat, prop)

myspecies <- "WHB"
mystock   <- "whb-comb"

# list the available files within the directory
files.list <- list.files(path=datapath, 
                         pattern=paste("WGWIDE catchesbyrect", toupper(myspecies)),
                         full.names=TRUE )

# read the files
# f <- files.list[1]

catch_by_species_year_country_raw <- data.frame(stringsAsFactors = FALSE)

for (f in files.list) {
  print(f)
  catch_by_species_year_country_raw <-
    bind_rows(
      catch_by_species_year_country_raw,
      read_excel(f, col_names=TRUE, col_types="text") %>% 
        lowcase() %>% 
        dplyr::mutate(across(c("catch", "lat","lon"), as.numeric ))  %>%
        dplyr::mutate(across(c("year"), as.integer)) %>% 
        dplyr::mutate(pnum = as.integer(gsub("quarter","", pnum))) %>% 
        dplyr::mutate(ptype = toupper(ptype)) %>% 
        
        filter(catch >0) %>% 
        
        dplyr::select(-lat, -lon) %>% 
        dplyr::left_join(icesrect, by=c("rect")) %>% 
        dplyr::mutate(eez      = geo_inside(lon=lon, lat=lat, map=eez, variable="Territory1")) %>%
        
        # replace country names
        dplyr::mutate(country = ifelse(country %in% c("UKE","GBR.E") , "GBR.EW", country)) %>% 
        dplyr::mutate(country = ifelse(country %in% c("IMN","GUY","JEY"), 
                                       paste("UK",country,sep="."), 
                                       country)) %>% 
        dplyr::mutate(country = ifelse(country %in% c("BES","SPA")      , "ESP", country)) %>% 
        dplyr::mutate(country = ifelse(country %in% c("POR")            , "PRT", country)) %>% 
        dplyr::mutate(country = ifelse(country %in% c("GER")            , "DEU", country)) %>% 
        dplyr::mutate(country = ifelse(country %in% c("IRL","ESP","NLD","DNK","DEU","FRA",
                                                      "FRA","PRT","SWE","POL","LTU","EST","BEL"), 
                                     paste("EU",country,sep="."), 
                                     country)) %>% 
      dplyr::mutate(country = gsub("^GBR","UK",country))

    )  
}


catch_by_species_year_country_raw_zbyrect <-
  catch_by_species_year_country_raw %>% 
  dplyr::select(-eez) %>% 
  left_join(dplyr::select(zbyrect,
                          rect, zone, prop), 
            by="rect") %>% 
  mutate(catch = catch * prop/100)

# catch by year
catch_by_species_year <- 
  catch_by_species_year_country_raw %>% 
  group_by(species, year, rect, lat, lon) %>% 
  summarize(catch = sum(catch, na.rm=TRUE)) %>% 
  
  mutate(catch_interval = cut(as.integer(catch), breaks=c(1 %o% 10^(0:12)), dig.lab=10 ) ) %>% 
  mutate(catch_interval = gsub(" ","", catch_interval)) %>% 
  filter(!is.na(catch_interval)) %>% 
  filter(grepl("\\,", catch_interval)) %>%   
  group_by(species, year, rect, lat, lon, catch_interval) %>% 
  summarize(catch = sum(catch, na.rm=TRUE))



# su(catch_by_species_year_country_raw$country)
# catch_by_species_year_country_raw %>% filter(grepl("IMN", country)) %>% View()

# mapplots::ices.rect(catch_by_species_year_country_raw$rect) %>% View()

# -------------------------------------------------------------------------------------
# Now read the new catch by rectangle dataset
# -------------------------------------------------------------------------------------

catch_by_rect2 <-
  read.csv(file.path(datapath2, "combined_catch_corrected.csv"), header=TRUE) %>% 
  lowcase() %>% 
  dplyr::rename(pnum = month) %>% 
  dplyr::mutate(ptype = "M") %>% 
  dplyr::rename(rect = icesrect) %>%       
  filter(catch >0) %>% 
  
  dplyr::left_join(icesrect, by=c("rect")) %>% 
  dplyr::mutate(eez      = geo_inside(lon=lon, lat=lat, map=eez, variable="Territory1")) %>%
  
  # replace country names
  # dplyr::mutate(country = ifelse(country %in% c("UKE","GBR.E") , "GBR.EW", country)) %>% 
  # dplyr::mutate(country = ifelse(country %in% c("IMN","GUY","JEY"), 
  #                                paste("UK",country,sep="."), 
  #                                country)) %>% 
  dplyr::mutate(country = ifelse(country %in% c("BES","SPA")      , "ESP", country)) %>% 
  dplyr::mutate(country = ifelse(country %in% c("POR")            , "PRT", country)) %>% 
  dplyr::mutate(country = ifelse(country %in% c("GER")            , "DEU", country)) %>% 
  dplyr::mutate(country = ifelse(country %in% c("IRL","ESP","NLD","DNK","DEU","FRA",
                                                "FRA","PRT","SWE","POL","LTU","EST","BEL"), 
                               paste("EU",country,sep="."), 
                               country)) %>% 
  dplyr::mutate(country = ifelse(grepl("^GBR|^UK|^IMN|^GUY|^JEY",country), "GBR", country)) %>% 

  dplyr::mutate(zone = ifelse(grepl("^int|^xin",tolower(zone)) & lat >= 62             , "INN", zone)) %>% 
  dplyr::mutate(zone = ifelse(grepl("^int|^xin",tolower(zone)) & lat < 62              , "INW", zone)) %>% 
  dplyr::mutate(zone = ifelse(grepl("faroes",tolower(zone))                            , "FRO", zone)) %>% 
  dplyr::mutate(zone = ifelse(grepl("norway",tolower(zone))                            , "NOR", zone)) %>% 
  dplyr::mutate(zone = ifelse(grepl("iceland",tolower(zone))                           , "ISL", zone)) %>% 
  dplyr::mutate(zone = ifelse(grepl("greenland",tolower(zone))                         , "GRL", zone)) %>% 
  dplyr::mutate(zone = ifelse(grepl("jersey",tolower(zone))                            , "GBR", zone)) %>% 
  dplyr::mutate(zone = ifelse(grepl("svalbard",tolower(zone))                          , "SVA", zone)) %>% 
  dplyr::mutate(zone = ifelse(grepl("jan mayen",tolower(zone))                         , "SJM", zone)) %>% 
  dplyr::mutate(zone = ifelse(grepl("^eu",tolower(zone)) & eez %in% c("United Kingdom"), "GBR", zone)) %>% 
  dplyr::mutate(zone = ifelse(grepl("^eu",tolower(zone))                               , "EU27", zone)) %>% 
  dplyr::mutate(zone = ifelse(grepl("^eu",tolower(zone)) & as.integer(lat)>=60         , "GBR", zone)) %>% 
  dplyr::mutate(zone = ifelse(grepl("^uk",tolower(zone))                               , "GBR", zone)) %>% 
  dplyr::mutate(zone = ifelse(grepl("^special",tolower(zone))                          , "GBR_FRO", zone)) %>% 
  dplyr::mutate(zone = ifelse(grepl("^unk",tolower(zone))                              , NA, zone)) 

# catch_by_rect2 %>% distinct(zone) %>% View()
# catch_by_rect2 %>% distinct(country) %>% View()
# catch_by_rect2 %>% filter(zone=="EU27", lat >= 60) %>% View()

# catch_by_rect2 %>%
#   rename(month=pnum) %>%
#   dplyr::select(-eez, -ptype, -h, -i) %>%
#   write.csv(., file=file.path(datapath2, "combined_catch_corrected_zones.csv"), row.names = FALSE)

catch_by_rect2_zbyrect <-
  catch_by_rect2 %>% 
  dplyr::select(-eez) %>% 
  left_join(dplyr::select(zbyrect,
                          rect, zone, prop), 
            by="rect") %>% 
  mutate(catch = catch * prop)


compare <- 
  iAdvice %>% 
  filter(stockkeylabelold==mystock) %>% 
  filter(adviceonstock==TRUE) %>% 
  dplyr::select(year=tacyear, advice=advisedlandingsmax, landings_ices=landings, catch_ices=catches) %>% 
  left_join(
    catch_by_species_year %>% 
      group_by(year) %>% 
      summarise(catch_byrect1=sum(catch, na.rm=TRUE)), 
    by="year") %>% 
  left_join(
    catch_by_rect2 %>% 
      group_by(year) %>% 
      summarise(catch_byrect2=sum(catch, na.rm=TRUE)), 
    by="year") %>% 
  filter(year >= 2011, year <= 2020) %>% 
  mutate(perc1 = scales::percent(catch_byrect1 / catch_ices-1, accuracy=1)) %>% 
  mutate(perc2 = scales::percent(catch_byrect2 / catch_ices-1, accuracy=1)) 

writexl::write_xlsx(compare, path="compare1.xlsx")
  

# compare by country, year and source
compare2 <- 
  bind_rows(
    catch_by_species_year_country_raw %>% 
      mutate(country = ifelse(grepl("^UK",country),"UK",country)) %>% 
      group_by(year, country) %>% 
      summarise(catch=sum(catch, na.rm=TRUE)) %>% 
      mutate(source="wgwide"),
    catch_by_rect2 %>% 
      group_by(year, country) %>% 
      summarise(catch=sum(catch, na.rm=TRUE)) %>% 
      mutate(source="datacall")
  ) %>% 
  filter(year >= 2011, year <= 2020)

compare2 %>% 
  reshape2::dcast(country ~ year+source, sum, value.var="catch", margins="country") %>% 
  writexl::write_xlsx(., path="compare2.xlsx")

# compare catches by zone from allocations in data call and calculated from zbyrect
compare3 <- 
  bind_rows(
    catch_by_rect2_zbyrect %>% 
      rename(zone=zone.x) %>% 
      group_by(year, zone) %>% 
      summarise(catch=sum(catch,na.rm=TRUE)) %>% 
      mutate(source="allocated"),
    catch_by_rect2_zbyrect %>% 
      rename(zone=zone.y) %>% 
      group_by(year, zone) %>% 
      summarise(catch=sum(catch,na.rm=TRUE)) %>% 
      mutate(source="calculated")
  ) %>% 
  mutate(zone = ifelse(zone == "EU"     , "EU27", zone)) %>% 
  mutate(zone = ifelse(zone == "UK"     , "GBR", zone)) %>% 
  mutate(zone = ifelse(zone == "Faroes"   , "FRO", zone)) %>% 
  mutate(zone = ifelse(zone == "Greenland", "GRL", zone)) %>% 
  mutate(zone = ifelse(zone == "Iceland"  , "ISL", zone)) %>% 
  mutate(zone = ifelse(zone == "Norway"  , "NOR", zone)) %>% 
  mutate(zone = ifelse(zone == "Inter. west"  , "INT", zone)) %>% 
  mutate(zone = ifelse(zone == "Inter. West"  , "INT", zone)) %>% 
  mutate(zone = ifelse(zone == "Inter. Norw. Sea"  , "INT", zone)) %>% 
  mutate(zone = ifelse(zone == "Svalbard"  , "SVA", zone)) %>% 
  mutate(zone = ifelse(zone == "Special area UK/FO"  , "GBR_FRO", zone)) %>% 
  mutate(zone = ifelse(zone == "XIN"     , "INT", zone)) %>% 
  mutate(zone = ifelse(zone == "Jersey"     , "GBR", zone)) %>% 
  mutate(zone = ifelse(zone == "Jan Mayen"  , "SJM", zone)) %>% 
  
  group_by(year, zone, source) %>% 
  summarise(catch=sum(catch, na.rm=TRUE))


# su(compare3$zone)

# compare <- 
#   iAdvice %>% 
#   filter(stockkeylabelold==mystock) %>% 
#   filter(adviceonstock==TRUE) %>% 
#   dplyr::select(year=tacyear, advice=advisedlandingsmax, officiallandings, landings_ices=landings, catch_ices=catches) %>% 
#   left_join(catch_by_species_year %>% 
#               group_by(year) %>% 
#               summarise(catch_byrect=sum(catch, na.rm=TRUE)), 
#             by="year") %>% 
#   
#   drop_na(catch_byrect) %>% 
#   
#   mutate(prop_catch = catch_byrect/catch_ices) %>% 
#   mutate(prop_catch2 = scales::percent(prop_catch)) %>% 
#   
#   mutate(prop_landings = catch_byrect/landings_ices) %>% 
#   mutate(prop_landings2 = scales::percent(prop_landings))


# ices <-  
#   read_excel("C:/DATA/Onedrive - PFA/Documents/iWGWIDE/2021/06. Data/mac.27.nea/ICES MAC catch by country.xlsx",
#              col_names=TRUE, col_types="text", sheet="Sheet2") %>%
#   lowcase() %>% 
#   mutate(proportionlandings = as.numeric(proportionlandings)) %>% 
#   mutate(subtotal           = as.numeric(subtotal)) %>% 
#   rename(year = country) %>% 
#   mutate(year = as.numeric(year))




```

**Comparison of updated catch by rectangle data**

Martin Pastoors, `r format(Sys.time(), '%d/%m/%Y')`


Comparison of ICES estimate of total catch with the summed catch by rectangle data from WGWIDE 2021 and from the Coastal States expert group data call. 

```{r, echo=FALSE, fig.asp=0.6, fig.align="center", message=FALSE, warning=FALSE}

compare %>% 
  dplyr::select(year, catch_ices, catchbyrect_wgwide=catch_byrect1, catchbyrect_datacall=catch_byrect2) %>% 
  pivot_longer(names_to = "var", values_to = "data", "catch_ices":"catchbyrect_datacall") %>% 
  mutate(var = factor(var, levels=c("catch_ices","catchbyrect_wgwide", "catchbyrect_datacall"))) %>% 
  
  ggplot(aes(x=year, y=data)) +
  theme_publication() +
  geom_point(aes(colour=var)) +
  geom_line(aes(colour=var)) +
  expand_limits(y=0) +
  scale_x_continuous(breaks=2011:2020) +
  labs (x="", y="tonnes", title="WHB catch (tonnes)", colour="")


```

\newpage

Comparison of ICES estimate of total catch with the summed catch by rectangle data from WGWIDE 2021 and from the Coastal States expert group data call. 

```{r, echo=FALSE, fig.asp=1.4, fig.align="center", message=FALSE, warning=FALSE}

compare2 %>% 
  filter(!grepl("^EU", country)) %>% 
  ggplot(aes(x=year, y=catch)) +
  theme_publication() +
  theme(strip.background = element_rect(colour="black", size =0.2),
        plot.margin      = unit(c(0,0,0,0),"cm")) +
  # geom_point(aes(colour=source)) +
  # geom_line(aes(colour=source)) +
  geom_bar(aes(fill=source), stat="identity", position=position_dodge2(width = 0.9, preserve = "single")) + 
  expand_limits(y=0) +
  scale_x_continuous(breaks=scales::pretty_breaks()) +
  # scale_fill_manual(palette=viridis_pal(option = "plasma")(10)[c(1,4)]) +
  labs (x="", y="tonnes", title="WHB catch (tonnes)", colour="") +
  facet_wrap(~country)

# scales::show_col(scales::viridis_pal(option = "plasma")(10)[c(1,4)])

  
```

\newpage

Comparison of catch by rectangle (summed) by country for EU27 countries. Note that Poland is not included in the new data from the data call. 

```{r, echo=FALSE, fig.asp=1.4, fig.align="center", message=FALSE, warning=FALSE}

compare2 %>% 
  filter(grepl("^EU", country)) %>% 
  ggplot(aes(x=year, y=catch)) +
  theme_publication() +
  theme(strip.background = element_rect(colour="black", size =0.2),
        plot.margin      = unit(c(0,0,0,0),"cm")) +
  # geom_point(aes(colour=source)) +
  # geom_line(aes(colour=source)) +
  geom_bar(aes(fill=source), stat="identity", position=position_dodge2(width = 0.9, preserve = "single")) + 
  expand_limits(y=0) +
  scale_x_continuous(breaks=scales::pretty_breaks()) +
  labs (x="", y="tonnes", title="WHB catch (tonnes)", colour="") +
  facet_wrap(~country, ncol=3)


```

\newpage

Catches by zone, either from allocation or calculated from zbyrect 

```{r, echo=FALSE, fig.asp=1.4, fig.align="center", message=FALSE, warning=FALSE}

compare3 %>% 
  drop_na(zone) %>% 
  ggplot(aes(x=year, y=catch)) +
  theme_publication() +
  theme(strip.background = element_rect(colour="black", size =0.2),
        plot.margin      = unit(c(0,0,0,0),"cm")) +
  # geom_point(aes(colour=source)) +
  # geom_line(aes(colour=source)) +
  geom_bar(aes(fill=source), stat="identity", position=position_dodge2(width = 0.9, preserve = "single")) + 
  
  expand_limits(y=0) +
  scale_x_continuous(breaks=scales::pretty_breaks()) +
  labs (x="", y="tonnes", title="WHB catch (tonnes)", colour="") +
  facet_wrap(~zone, ncol=3)


# compare3 %>% 
#   filter(source=="allocated") %>% 
#   group_by(zone) %>% 
#   summarise(catch=sum(catch, na.rm=TRUE)) %>% 
#   group_by() %>% 
#   mutate(prop = catch/sum(catch)) %>% 
#   View()


```

\newpage

Catches by zone (presence-absence) - all years combined

```{r, echo=FALSE, fig.asp=1.4, fig.align="center", message=FALSE, warning=FALSE}

catch_by_rect2 %>%
  # filter(zone=="EU27") %>% 
  drop_na(zone) %>% 
  group_by(zone, lat, lon) %>% 
  summarise(catch=sum(catch, na.rm=TRUE)) %>% 
  
  ggplot(aes(x=lon, y=lat)) +
  theme_publication() + theme(legend.position="none") +
  coord_quickmap(xlim=c(-40,20), ylim=c(45,75)) +
  # coord_quickmap(xlim=c(-20,10), ylim=c(45,65)) +
  geom_polygon(data=world.df, aes(long, lat, group = group), fill = "gray") +
  geom_polygon(data=eez.df, aes(long, lat, group = group), fill = NA, colour="gray") +
  geom_rect(aes(xmin=lon-0.5, xmax=lon+0.5, ymin=lat-0.25, ymax=lat+0.25, label=zone, fill=zone), alpha=0.5) +
  facet_wrap(~zone, ncol=3)

```

\newpage

Catches by zone (catch amounts) - all years combined

```{r, echo=FALSE, fig.asp=1.4, fig.align="center", message=FALSE, warning=FALSE}

catch_by_rect2 %>%
  # filter(zone=="EU27") %>% 
  drop_na(zone) %>% 
  group_by(zone, lat, lon) %>% 
  summarise(catch=sum(catch, na.rm=TRUE)) %>% 
  
  ggplot(aes(x=lon, y=lat)) +
  theme_publication() + theme(legend.position="none") +
  coord_quickmap(xlim=c(-40,20), ylim=c(45,75)) +
  # coord_quickmap(xlim=c(-20,10), ylim=c(45,65)) +
  geom_polygon(data=world.df, aes(long, lat, group = group), fill = "gray") +
  geom_polygon(data=eez.df, aes(long, lat, group = group), fill = NA, colour="gray") +
  geom_point(aes(fill=zone, colour=zone, size=catch), shape=21, alpha=0.5) +
  scale_size_continuous(range=c(0.1,10)) +
  facet_wrap(~zone, ncol=3)


```
