---
output: 
  word_document:
    reference_docx: ../report_template_v1.5.dotx
---

```{r setup, include=FALSE}

# =======================================================================================
# Catch by rectangle tester.Rmd
# 
# Generate catch by rectangle for test dataset
#
# 04/02/2021 Adapted from blue whiting code; made more generic
# 01/03/2022 Adapted the code for reading / plotting single dataset
# =======================================================================================

require("knitr")
knitr::opts_chunk$set(echo = FALSE,	message = FALSE,	warning = FALSE,	comment = "",	crop = TRUE )
knitr::opts_chunk$set(fig.width=10) 

rm(list=ls())

library(tidyverse)    # tidying packages
library(readxl)       # read excel
library(lubridate)
library(sf)
library(sp)
library(RColorBrewer)
library(pander)

library(maps)
library(lubridate)
library(viridis)

# source my utils
source("../../prf/R/my utils.r")
source("../../gisland/r/geo_inside.R")

dropboxdir <- paste(get_dropbox(), "/iAdvice", sep="")
onedrive   <- get_onedrive()

load(file.path(onedrive, "rdata/world.df.RData"))
load(file.path(onedrive, "rdata/eez.df.RData"))
load(file.path(onedrive, "rdata/fao.df.RData"))
load(file.path(onedrive, "rdata/icesrectangles.df.RData"))

ices2 <- icesrectangles.df %>% group_by(rect) %>% filter(row_number()==1) %>% 
  mutate(
    lon=(EAST+WEST)/2,
    lat=(SOUTH+NORTH)/2
  ) %>% 
  dplyr::select(rect, lon, lat)


# Data path and file
datapath <- "C:/DATA/Onedrive - PFA/Documents/CS WHB/Data/"
myfile  <- file.path(datapath, "Catch data/WHB_NLD.csv") 

# read raw data file and do conversions
catch_by_rect <- 
  read.csv(file=myfile, header=TRUE, colClasses = "character", stringsAsFactors = FALSE) %>% 
  lowcase() %>% 
  rename(rect = icesrect) %>% 
  dplyr::mutate(across(c("catch"), as.numeric ))  %>%
  dplyr::mutate(across(c("year", "month"), as.integer)) %>% 

  filter(catch >0) %>% 
  
  dplyr::left_join(ices2, by=c("rect")) 

zbycoastalstate <- 
  readxl::read_xlsx(path=file.path(datapath, "ZbyCoastalstate_withUK.xlsx"),
                    col_types="text",
                    .name_repair = "unique") %>% 
  lowcase() %>% 
  mutate(across(c("lat","lon","area", "prop"), as.numeric)) %>% 
  dplyr::select(-geometry)


plot_catch_by_year <- function(myspecies="MAC", myyears=NA, plot_catch=TRUE, plot_ssb=FALSE, ncol=6, xlim=NA, ylim=NA) {
  
  catch2 <-
    catch_by_species_year %>% 
    filter(species == myspecies) %>% 
    {if ((length(myyears) == 1) & all(is.na(myyears))) {
      filter(.)
    } else {
      filter(., year %in% myyears)
    }} %>% 
    
    {if (!all(is.na(myyears))) bind_rows(., data.frame(species=myspecies, year = myyears, catch_interval="(1,10]", catch=0)) else (.) }

  if(all(is.na(xlim))) {xlim <- range(catch2$lon, na.rm=TRUE)}
  if(all(is.na(ylim))) {ylim <- range(catch2$lat, na.rm=TRUE)}

  tc <-
    catch2 %>% 
    group_by(species, year) %>% 
    summarize(catch = sum(catch, na.rm=TRUE)) %>% 
    mutate(catch  = as.integer(catch) ) %>% 
    
    group_by(species) %>% 
    mutate(catch2 = ylim[1] + (catch / max(catch, na.rm=TRUE) * (ylim[2] - ylim[1])) )
  
  ggplot2::map_data("world", xlim = xlim, ylim = ylim) %>% 
    ggplot(aes(long, lat)) +
    geom_polygon(aes(long, lat, group = group), fill = "gray") +
    coord_quickmap(xlim = xlim, ylim = ylim, expand = FALSE) +
    
    theme_publication() +
    theme(panel.border     = element_rect(colour="black" , size=0.2),
          panel.grid.major = element_blank(),
          strip.background = element_rect(colour="black", size =0.2),
          plot.margin      = unit(c(0,0,0,0),"cm"),
          plot.title       = element_text(hjust=0, vjust=0, size=10),
          axis.text        = element_text(size=6),
          legend.key.width = unit(0.4, "cm"), 
          axis.title       = element_blank()) +
    
    geom_tile(data=catch2, aes(x=lon, y=lat, fill = catch_interval, height=0.5, width=1), colour=NA, alpha=1.0) +
    scale_fill_viridis(option = "plasma", direction = -1, discrete=TRUE) +
    
    geom_label(data=tc, aes(label=paste0("C ",catch)), x=-Inf, y=Inf, hjust=0, vjust=1, inherit.aes=FALSE) +
  
    labs(x = NULL, y = NULL, size = "tons", title=unique(catch2$species)) +
    guides(fill = guide_legend(nrow = 1, title="catch (tonnes)")) +
    facet_wrap( ~ year, ncol=ncol)

} # end of plot_catch_by_year



```


**Catch by rectangle tester: `r myfile`**

Martin Pastoors, `r format(Sys.time(), '%d/%m/%Y')`

Catch by rectangle from `r myfile`. 

```{r echo=FALSE, fig.align="center", fig.asp=1.2, message=FALSE, warning=FALSE}

t <-
  catch_by_rect %>% 
  group_by(species, year, lat, lon) %>% 
  summarize(catch = as.integer(sum(catch, na.rm=TRUE))) %>% 
  mutate(catch2 = cut(catch, breaks=c(0,100,1000,5000,25000,500000),  dig.lab=10))

tt <-
  t %>% 
  group_by(species, year) %>% 
  summarize(catch = as.integer(sum(catch, na.rm=TRUE))) 
  
# unique(t$catch2)

xlim <- range(t$lon, na.rm=TRUE)
ylim <- range(t$lat, na.rm=TRUE)

ggplot2::map_data("world", xlim = xlim, ylim = ylim) %>% 
  ggplot(aes(long, lat)) +
  
  theme_publication() +
  theme(panel.border     = element_rect(colour="black" , size=0.2),
        panel.grid.major = element_blank(),
        strip.background = element_rect(colour="black", size =0.2),
        plot.margin      = unit(c(0,0,0,0),"cm"),
        plot.title       = element_text(hjust=0, vjust=0, size=10),
        axis.text        = element_text(size=6),
        legend.key.width = unit(0.4, "cm"), 
        axis.title       = element_blank()) +
  
  coord_quickmap(xlim = xlim, ylim = ylim, expand = FALSE) +
  
  geom_tile(data=t, aes(x=lon, y=lat, fill = catch2, height=0.5, width=1), colour=NA, alpha=1.0) +
  scale_fill_viridis(option = "plasma", direction = -1, discrete=TRUE) +
  
  geom_polygon(aes(long, lat, group = group), fill = "gray") +
  geom_polygon(data=eez.df, aes(long, lat, group = group), fill = NA, colour="gray") +
  
  geom_label(data=tt, aes(label=catch), 
             x=Inf, y=-Inf, hjust=1, vjust=0, inherit.aes=FALSE, fill="white", alpha=0.8) +

  labs(x = NULL, y = NULL, size = "tons", title=myfile) +
  guides(fill = guide_legend(nrow = 1, title="catch (tonnes)")) +
  facet_wrap( ~ year, ncol=4)



```

\newpage

Catch by rectangle and zone from `r myfile`. 

```{r echo=FALSE, fig.align="center", fig.asp=1.2, message=FALSE, warning=FALSE}

t <-
  catch_by_rect %>% 
  group_by(species, zone, year, lat, lon) %>% 
  summarize(catch = as.integer(sum(catch, na.rm=TRUE))) %>% 
  mutate(catch2 = cut(catch, breaks=c(0,100,1000,5000,25000,500000),  dig.lab=10))

tt <-
  t %>% 
  group_by(species, year) %>% 
  summarize(catch = as.integer(sum(catch, na.rm=TRUE))) 
  
# unique(t$catch2)

xlim <- range(t$lon, na.rm=TRUE)
ylim <- range(t$lat, na.rm=TRUE)

ggplot2::map_data("world", xlim = xlim, ylim = ylim) %>% 
  ggplot(aes(long, lat)) +
  
  theme_publication() +
  theme(panel.border     = element_rect(colour="black" , size=0.2),
        panel.grid.major = element_blank(),
        strip.background = element_rect(colour="black", size =0.2),
        plot.margin      = unit(c(0,0,0,0),"cm"),
        plot.title       = element_text(hjust=0, vjust=0, size=10),
        axis.text        = element_text(size=6),
        legend.key.width = unit(0.4, "cm"), 
        axis.title       = element_blank()) +
  
  coord_quickmap(xlim = xlim, ylim = ylim, expand = FALSE) +
  
  geom_point(data=t, aes(x=lon, y=lat, size = catch, fill=zone), colour="gray", alpha=0.8, shape=21) +

  geom_polygon(aes(long, lat, group = group), fill = "gray") +
  geom_polygon(data=eez.df, aes(long, lat, group = group), fill = NA, colour="gray") +
  
  geom_label(data=tt, aes(label=catch), 
             x=Inf, y=-Inf, hjust=1, vjust=0, inherit.aes=FALSE, fill="white", alpha=0.8) +

  labs(x = NULL, y = NULL, size = "tons", title=myfile) +
  guides(fill = guide_legend(nrow = 1, title="catch (tonnes)")) +
  facet_wrap( ~ year, ncol=4)




```

\newpage

Catch by rectangle and zone from `r myfile` in the area between Faroer and GBR. Based on Dutch data submission. 

```{r echo=FALSE, fig.align="center", fig.asp=1.2, message=FALSE, warning=FALSE}

t <-
  catch_by_rect %>% 
  filter(lon >= -13, lon <= -4) %>% 
  filter(lat >= 58, lat <= 61) %>% 
  group_by(species, zone, rect, year, lat, lon) %>% 
  summarize(catch = as.integer(sum(catch, na.rm=TRUE))) %>% 
  group_by(species, rect, year, lat, lon) %>% 
  mutate(value = catch /sum(catch,na.rm=TRUE)) %>% 
  mutate(allocation = "data")

tt <-
  t %>% 
  group_by(species, year) %>% 
  summarize(catch = as.integer(sum(catch, na.rm=TRUE))) 
  
# unique(t$catch2)

xlim <- range(t$lon, na.rm=TRUE)
ylim <- range(t$lat, na.rm=TRUE)

ggplot2::map_data("world", xlim = xlim, ylim = ylim) %>% 
  ggplot(aes(long, lat)) +
  
  theme_publication() +
  coord_equal(xlim = xlim, ylim = ylim, expand = FALSE) +
  
  scatterpie::geom_scatterpie(data=t,
                              aes(x=lon, y=lat, group=rect, r=0.2),
                              cols="zone", long_format = TRUE, pie_scale = 1, colour=NA)  +

  # geom_point(data=t, aes(x=lon, y=lat, size = value, fill=zone), colour="gray", alpha=0.8, shape=21) +

  geom_polygon(aes(long, lat, group = group), fill = "gray") +
  geom_polygon(data=icesrectangles.df, aes(long, lat, group = group), fill = NA, colour="gray") +
  geom_polygon(data=eez.df, aes(long, lat, group = group), fill = NA, colour="gray") +
  
  geom_label(data=tt, aes(label=catch), 
             x=Inf, y=-Inf, hjust=1, vjust=0, inherit.aes=FALSE, fill="white", alpha=0.8) +

  labs(x = NULL, y = NULL, size = "tons", title=myfile) +
  guides(fill = guide_legend(nrow = 1, title="catch (tonnes)")) +
  facet_wrap( ~ year, ncol=3)




```

\newpage

Catch by rectangle and zone from `r myfile` in the area between Faroer and GBR. Based on allocation from zone by rect.  

```{r echo=FALSE, fig.align="center", fig.asp=1.2, message=FALSE, warning=FALSE}

s <-
  catch_by_rect %>% 
  filter(lon >= -13, lon <= -4) %>% 
  filter(lat >= 58, lat <= 61) %>% 
  group_by(species, rect, year, lat, lon) %>% 
  summarize(catch = as.integer(sum(catch, na.rm=TRUE))) %>% 
  left_join(dplyr::select(zbycoastalstate,
                          rect, zone, prop)) %>% 
  mutate(catch = prop * catch) %>% 
  group_by(species, rect, year, lat, lon) %>%
  mutate(value = catch / sum(catch, na.rm=TRUE)) %>% 
  mutate(allocation = "zone_by_rect")

ss <-
  s %>% 
  group_by(species, year) %>% 
  summarize(catch = as.integer(sum(catch, na.rm=TRUE))) 
  
# unique(t$catch2)

xlim <- range(s$lon, na.rm=TRUE)
ylim <- range(s$lat, na.rm=TRUE)

ggplot2::map_data("world", xlim = xlim, ylim = ylim) %>% 
  ggplot(aes(long, lat)) +
  
  theme_publication() +
  coord_equal(xlim = xlim, ylim = ylim, expand = FALSE) +
  
  scatterpie::geom_scatterpie(data=s,
                              aes(x=lon, y=lat, group=rect, r=0.2),
                              cols="zone", long_format = TRUE, pie_scale = 1, colour=NA)  +

  # geom_point(data=t, aes(x=lon, y=lat, size = value, fill=zone), colour="gray", alpha=0.8, shape=21) +

  geom_polygon(aes(long, lat, group = group), fill = "gray") +
  geom_polygon(data=icesrectangles.df, aes(long, lat, group = group), fill = NA, colour="gray") +
  geom_polygon(data=eez.df, aes(long, lat, group = group), fill = NA, colour="gray") +
  
  geom_label(data=ss, aes(label=catch), 
             x=Inf, y=-Inf, hjust=1, vjust=0, inherit.aes=FALSE, fill="white", alpha=0.8) +

  labs(x = NULL, y = NULL, size = "tons", title=myfile) +
  guides(fill = guide_legend(nrow = 1, title="catch (tonnes)")) +
  facet_wrap( ~ year, ncol=3)



```

\newpage

Comparison of catch by rectangle and zone from `r myfile` in the area between Faroer and GBR. Based on logbook allocation or allocation from zone_by_rect.  

```{r echo=FALSE, fig.align="center", fig.asp=0.6, message=FALSE, warning=FALSE}

bind_rows(t,s) %>% 
  ggplot(aes(x=year, y=catch)) +
  theme_publication() +
  geom_bar(aes(fill=zone), stat="identity") +
  scale_x_continuous(breaks=seq(2010,2020,1)) +
  facet_wrap(~allocation) 

bind_rows(t,s) %>% 
  # mutate(value = scales::percent(value, accuracy=1)) %>% 
  reshape2::dcast(year ~ allocation + zone, value=value, mean) %>% 
  mutate(across(c(data_FRO, data_GBR, zone_by_rect_FRO, zone_by_rect_GBR, zone_by_rect_GBR_FRO), scales::percent)) %>% 
  pander::pandoc.table(
    style="simple",
    missing=" ",
    split.tables = 400
  )


```

