---
output: 
  word_document:
    reference_docx: ../report_template_v1.5.dotx
---

```{r setup, include=FALSE}

# =======================================================================================
# mackerel catch by rectangle.Rmd
# 
# Generate stock summary report for WGWIDE
#
# 03/09/2017 first version
# 06/08/2018 UPdated and prepared for WGWIDE 2018
# 07/09/2018 Added link to SAG database for plotting. Currently reading from downloaded data. 
#            Needs to be made through live connection. 
# 16/11/2021 Adapted for Coastal States meeting
# 17/01/2022 Adapted with corrected data series
# 02/10/2022 Plot of catch by zone; need to update the catch data in the first code section
# 01/03/2023 Adapted for Coastal States negotiations meeting
# 06/03/2023 Only CS WG data used
# =======================================================================================

require("knitr")
knitr::opts_chunk$set(echo = FALSE,	message = FALSE,	warning = FALSE,	comment = "",	crop = TRUE )
knitr::opts_chunk$set(fig.width=10) 

rm(list=ls())

library(tidyverse)    # tidying packages
library(readxl)       # read excel
library(lubridate)
library(sf)
library(sp)
library(RColorBrewer)
library(pander)

# library(gisland)      # devtools::install_github("einarhjorleifsson/gisland", dependencies = TRUE)

library(maps)
# library(mapdata)
# library(mapproj)
library(lubridate)
library(viridis)
library(scatterpie)  # plotting multiple pie charts on a map

# install.packages("mregions")
# library(mregions)

# library(animation)
# library(gganimate)

# source my utils
source("../../prf/R/my utils.r")
source("../../prf/R/geo_inside.R")

dropboxdir <- paste(get_dropbox(), "/iAdvice", sep="")

# Load dataset
iAdvice <-
  loadRData(file=paste(dropboxdir, "/rdata/iAdvice.RData",sep="")) %>% 
  filter(stockkeylabelold == "mac-nea") %>% 
  filter(adviceonstock==TRUE)

# onedrive <- get_onedrive()
onedrive <- "C:/DATA/RDATA"

load(file.path(onedrive, "world_mr_df.RData"))
load(file.path(onedrive, "eez.df.RData"))
load(file.path(onedrive, "fao.df.RData"))
load(file.path(onedrive, "depth200.df.RData"))
load(file.path(onedrive, "rect_df.RData"))

load(file.path(onedrive, "eez.RData"))
load(file.path(onedrive, "fao.RData"))

load(file.path(onedrive, "rect_sf.RData"))
# load(file.path(onedrive, "ices.eez.sf.RData"))
load(file.path(onedrive, "eez.sf.RData"))
# load(file.path(onedrive, "rdata/eez.iho.sf.RData"))

# eez.sf <- mregions::mr_shp(key = "MarineRegions:eez") %>% sf::st_as_sf(res2)

icesrect <- 
  rect_df %>% 
  rename(rect=ICESNAME) %>% 
  group_by(rect) %>% 
  distinct(rect, .keep_all = TRUE) %>% 
  mutate(
    long = (WEST + EAST)/2,
    lat  = (SOUTH + NORTH)/2
  )
# write.csv(icesrect, file="icesrect.csv", row.names = FALSE)

# Data path
# datapath <- "C:/Users/Martin//Onedrive - PFA/Documents/iWGWIDE/2021/06. Data/_catch_by_rectangle"

# documentpath <- "D:/COASTALSTATES/CS MAC"
documentpath <- "C:/DATA/CS MAC"
datapath     <- file.path(documentpath, "Data", "Catch data")

# CS WG file
files.list <- file.path(datapath, "New-WGWIDE catchesbyrect MAC.xlsx")

# WGWIDE 2022 file
# files.list <- file.path(datapath, "WGWIDE catchesbyrect MAC.xlsx")

zbyrect <- 
  readxl::read_excel(
    path = file.path(documentpath,"Data", "ZbyCoastalstate_withUK.xlsx")
  ) %>% 
  dplyr::select(-geometry)

myspecies <- "MAC"
mystock   <- "mac-nea"

# list the available files within the directory
# files.list <- list.files(path=datapath, 
#                          pattern=paste("WGWIDE catchesbyrect", toupper(myspecies)),
#                          full.names=TRUE )

# read the files
# f <- files.list[1]

# russian_catch <-
#   read_excel(path=file.path(documentpath,"Data","Catch data", "Russian catch 2021.xlsx"), 
#              col_names=TRUE) 
  

catch_by_species_year_country_raw <- data.frame(stringsAsFactors = FALSE)

for (f in files.list) {
  print(f)
  catch_by_species_year_country_raw <-
    bind_rows(
      catch_by_species_year_country_raw,
      read_excel(f, col_names=TRUE, col_types="text") %>% 
        lowcase() %>% 
        dplyr::mutate(across(c("catch", "lat","lon"), as.numeric ))  %>%
        dplyr::mutate(across(c("year"), as.integer)) %>% 
        dplyr::mutate(pnum = as.integer(gsub("quarter","", pnum))) %>% 
        dplyr::mutate(ptype = toupper(ptype)) %>% 
        
        # change month to quarter
        # dplyr::mutate(pnum  = ifelse(ptype=="M", ceiling(as.numeric(pnum)/3), pnum)) %>% 
        # dplyr::mutate(ptype = ifelse(ptype=="M", "Q", ptype)) %>% 
        
        filter(catch >0) 
    )  
}


catch_by_species_year_country_raw <-
  catch_by_species_year_country_raw %>% 

  # replace lat long with values from icesrect (bottom-left points of rectangles)
  dplyr::select(-lat, -lon) %>% 
  dplyr::left_join(icesrect, by=c("rect")) %>% 
  rename(lon=long) %>% 
  
  # dplyr::filter((species == "hom" & lon > -25) | (species != "hom")) %>% 
  ungroup() %>% 
  drop_na(lat, lon) %>% 
  
  # No. Don't allocate eez here; use the zbyrect version for that
  # dplyr::mutate(eez      = geo_inside(lon=lon, lat=lat, map=eez, variable="Territory1")) %>%
  
  # ungroup() %>% 
  # dplyr::mutate(eez2     = geo_inside(lon=lon, lat=lat, map=eez, variable="Territory2")) %>%
  
  # replace country names
  dplyr::mutate(country = ifelse(country %in% c("UKE","GBR.E") , "GBR.EW", country)) %>% 
  dplyr::mutate(country = ifelse(country %in% c("IMN","GUY","JEY", "GGY"), 
                                 paste("GBR",country,sep="."), 
                                 country)) %>% 
  dplyr::mutate(country = ifelse(country %in% c("IMN")            , "GBR.IMN", country)) %>% 
  dplyr::mutate(country = ifelse(country %in% c("BES","SPA")      , "ESP", country)) %>% 
  dplyr::mutate(country = ifelse(country %in% c("POR")            , "PRT", country)) %>% 
  dplyr::mutate(country = ifelse(country %in% c("GER")            , "DEU", country)) %>% 
  
  dplyr::mutate(country = ifelse(grepl("GBR", country), "UK", country)) %>% 
  
  dplyr::mutate(
    region = case_when(
      country %in% c("IRL","ESP","NLD","DNK","DEU","FRA","PRT","SWE","POL","LTU","EST","BEL") ~ "EU27",
      TRUE ~ country
  ) )  
  # dplyr::bind_rows(dplyr::select(russian_catch,
  #                                year, species, region, catch))
  # 

catch_by_species_year_country_raw_zbyrect <-
  catch_by_species_year_country_raw %>% 
  filter(!(region=="RUS" & year==2021)) %>% 
  left_join(dplyr::select(zbyrect,
                          rect, eez=zone, prop), 
            by="rect") %>% 
  mutate(catch = catch * prop) %>% 
  mutate(eez = case_when(
    eez == "GBR"   ~ "UK",
    TRUE           ~ eez
  ))  
  # dplyr::bind_rows(russian_catch)


# unique(catch_by_species_year_country_raw_zbyrect$eez)
# catch_by_species_year_country_raw_zbyrect %>% 
#   filter(is.na(eez)) %>% 
#   View()

# su(catch_by_species_year_country_raw$country)
# catch_by_species_year_country_raw %>% filter(grepl("IMN", country)) %>% View()

# mapplots::ices.rect(catch_by_species_year_country_raw$rect) %>% View()

# catch by year
catch_by_species_year <- 
  catch_by_species_year_country_raw %>% 
  group_by(species, year, rect, lat, lon) %>% 
  summarize(catch = sum(catch, na.rm=TRUE)) %>% 
  
  mutate(catch_interval = cut(as.integer(catch), breaks=c(1 %o% 10^(0:12)), dig.lab=10 ) ) %>% 
  mutate(catch_interval = gsub(" ","", catch_interval)) %>% 
  filter(!is.na(catch_interval)) %>% 
  filter(grepl("\\,", catch_interval)) %>%   
  group_by(species, year, rect, lat, lon, catch_interval) %>% 
  summarize(catch = sum(catch, na.rm=TRUE))

   # map_data("worldHires", xlim = xlim, ylim = ylim) %>%

compare <- 
  iAdvice %>% 
  filter(stockkeylabelold==mystock) %>% 
  filter(adviceonstock==TRUE) %>% 
  dplyr::select(year=tacyear, advice=advisedlandingsmax, officiallandings, landings_ices=landings, catch_ices=catches) %>% 
  left_join(catch_by_species_year %>% 
              group_by(year) %>% 
              summarise(catch_byrect=sum(catch, na.rm=TRUE)), 
            by="year") %>% 
  mutate(prop_catch = catch_byrect/catch_ices) %>% 
  mutate(prop_catch2 = scales::percent(prop_catch)) %>% 
  
  mutate(prop_landings = catch_byrect/landings_ices) %>% 
  mutate(prop_landings2 = scales::percent(prop_landings))

# ices <-  
#   read_excel(file.path(documentpath, "2021/06. Data/mac.27.nea/ICES MAC catch by country.xlsx"),
#              col_names=TRUE, col_types="text", sheet="Sheet2") %>%
#   lowcase() %>% 
#   mutate(proportionlandings = as.numeric(proportionlandings)) %>% 
#   mutate(subtotal           = as.numeric(subtotal)) %>% 
#   rename(year = country) %>% 
#   mutate(year = as.numeric(year))

# ices2 <-
#   ices %>% 
#   dplyr::select(-misreported, -unallocated, -discards, -proportionlandings, -subtotal, -total) %>% 
#   tidyr::pivot_longer(names_to = "country", values_to = "catch", "belgium":"ussr/russianfederation") %>% 
#   drop_na(catch) %>% 
#   mutate(
#     catch = as.numeric(catch),
#     country = ifelse(grepl("faroe", country), "FRO", country),
#     country = ifelse(grepl("northern|jersey|isle of|guernsey|england|united", country), "UK", country),
#     country = ifelse(grepl("spain", country), "ESP", country),
#     country = ifelse(grepl("ussr", country), "RUS", country),
#     country = ifelse(grepl("germany", country), "DEU", country), 
#     country = ifelse(grepl("neth", country), "NLD", country), 
#     country = ifelse(grepl("portu", country), "PRT", country), 
#     country = ifelse(grepl("lith", country), "LTU", country), 
#     country = ifelse(grepl("belgium", country), "BEL", country), 
#     country = ifelse(grepl("france", country), "FRA", country), 
#     country = ifelse(grepl("denmark", country), "DNK", country), 
#     country = ifelse(grepl("sweden", country), "SWE", country), 
#     country = ifelse(grepl("ireland", country), "IRL", country), 
#     country = ifelse(grepl("norway", country), "NOR", country), 
#     country = ifelse(grepl("eston", country), "EST", country), 
#     country = ifelse(grepl("iceland", country), "ISL", country), 
#     country = ifelse(grepl("greenland", country), "GRL", country), 
#     country = ifelse(grepl("latvia", country), "LVA", country), 
#     country = ifelse(grepl("romania", country), "ROM", country), 
#     country = ifelse(grepl("poland", country), "POL", country), 
#    
#   ) %>% 
#   group_by(year, country) %>% 
#   summarise(catch = sum(catch, na.rm=TRUE)) %>% 
#   mutate(dataset="ICES2") %>% 
#   mutate(region = case_when(
#     country == "UK" ~ "UK",
#     country %in% c("BEL","DEU","DNK","NLD","FRA","POL","LTU","LVA","ESP","PRT","SWE","IRL","EST","ROM") ~ "EU27",
#     TRUE ~ country
#   ))

# catchcomp <-  
#   read_excel("C:/TEMP/MAC compare datasets2.xlsx",
#              col_names=TRUE, col_types="text", sheet="Sheet2") %>%
#   lowcase() %>% 
#   mutate(catch = as.numeric(catch)) %>% 
#   mutate(year  = as.integer(year)) %>% 
#   mutate(region = ifelse(country=="UK","UK","EU27")) %>% 
#   bind_rows(ices2) %>% 
#   bind_rows(
#     catch_by_species_year_country_raw %>% 
#       group_by(region, country, year) %>% 
#       summarise(catch = sum(catch, na.rm=TRUE)) %>% 
#       mutate(dataset = "ICESRECT")
#   )

# catchcomp %>% 
#   # group_by(dataset, year, region) %>% 
#   # summarise(catch=sum(catch,na.rm=TRUE)) %>% 
#   filter(dataset != "FDI2017") %>% 
#   filter(grepl("ICES", dataset)) %>% 
#   filter(catch > 10) %>% 
#   
#   ggplot(aes(x=year,y=catch,group=dataset)) +
#   theme_publication() +
#   geom_line(aes(colour=dataset)) +
#   geom_point(aes(colour=dataset)) +
#   expand_limits(y=0) +
#   facet_wrap(~country, scales="free_y")

# catchcomp %>% 
#   filter(dataset != "FDI2017") %>% 
#   filter(catch > 10) %>% 
#   pivot_wider(names_from = "dataset", values_from = "catch") %>% 
#   filter(!is.na(ACDR)) %>% 
#   mutate(
#     ices_diff = (ICES-ACDR)/ACDR,
#     ices2_diff = (ICES2-ACDR)/ACDR,
#     icesrect_diff = (ICESRECT-ACDR)/ACDR,
#     fdi2019_diff = (FDI2019-ACDR)/ACDR) %>% 
#   dplyr::select(-ACDR,-ICES,-FDI2019) %>% 
#   pivot_longer(names_to = "dataset", values_to = "diff", c("ices_diff","ices2_diff","icesrect_diff","fdi2019_diff")) %>% 
# 
#   ggplot(aes(x=year,y=diff,group=dataset)) +
#   theme_publication() +
#   geom_line(aes(colour=dataset)) +
#   geom_point(aes(colour=dataset)) +
#   geom_hline(yintercept=0) +
#   facet_wrap(~country)


```

**Catch history of mackerel**

Martin Pastoors, `r format(Sys.time(), '%d/%m/%Y')`

Catch overviews based on WGWIDE catch by rectangle data

Mackerel catch by Coastal State with the ICES advice superimposed

```{r, echo=FALSE, fig.asp=0.6, fig.align="center", message=FALSE, warning=FALSE}

# catch_by_species_year_country_raw %>% distinct(eez) %>% View()

t <-
  catch_by_species_year_country_raw %>% 
  group_by(species, year, region) %>% 
  summarize(catch = as.integer(sum(catch, na.rm=TRUE))) %>% 
  mutate(prop = catch / sum(catch, na.rm=TRUE))

tt <-
  iAdvice %>% 
  filter(tacyear %in% t$year) %>% 
  mutate(source="ICES") %>% 
  dplyr::select(year=tacyear, advice=advisedlandingsmax, source)

ttt <-
  t %>% 
  filter(year >= 2008) %>% 
  group_by(year) %>% 
  summarise(catch = sum(catch, na.rm=TRUE)) %>% 
  left_join(tt) %>% 
  mutate(overshoot = scales::percent(catch/advice-1, accuracy=1))


regions <- c("EU27","NOR","UK","FRO","RUS", "ISL","GRL")

t %>% 
  group_by(region) %>% 
  mutate(region=factor(region, levels=regions)) %>% 
  # filter(sum(prop) > 0.02) %>% 
  
  ggplot(aes(x=year,y=catch)) +
  theme_bw() +
  theme(legend.position="right") +
  geom_bar(aes(fill=region), 
                 stat="identity", position=position_stack(reverse=TRUE), colour="gray", linewidth=0.1) +
  geom_line(data=tt,
            aes(x=year,y=advice, colour=source), 
            linewidth=2) +
  geom_text(data=ttt,
            aes(x=year,y=catch, label=overshoot),
            vjust=0, nudge_y = 10000) +
  scale_color_manual(name='',
                     values=c('ICES'='black'))+
  
  geom_vline(xintercept=2013.5, linewidth=0.2, linetype="dashed") +
  geom_vline(xintercept=2020.5, linewidth=0.2, linetype="dashed") +

  labs(x="", y="tonnes", title="Mackerel catch by Party, ICES advice and overshoot", fill="Party") +
  scale_y_continuous(labels=scales::comma) +
  scale_fill_brewer(palette="Paired") +
  expand_limits(y=0)


```

_Figure 1: Mackerel catch by Party with the ICES advice superimposed._

\newpage

Mackerel catch proportions of the 'traditional' parties' catches

```{r, echo=FALSE, fig.asp=0.6, fig.align="center", message=FALSE, warning=FALSE}

# catch_by_species_year_country_raw %>% distinct(eez) %>% View()

regions <- c("EU27","NOR", "UK","FRO")

t <-
  catch_by_species_year_country_raw %>% 
  filter(region %in% regions) %>% 
  group_by(species, year, region) %>% 
  summarize(catch = as.integer(sum(catch, na.rm=TRUE))) %>% 
  mutate(prop = catch / sum(catch, na.rm=TRUE)) %>% 
  mutate(region=factor(region, levels=regions)) 
  
tt <-
  t %>% 
  mutate(region=factor(region, levels=regions)) %>% 
  filter(year %in% 2010:2020) %>% 
  group_by(region) %>% 
  summarise(catch = sum(catch)) %>% 
  ungroup() %>% 
  mutate(prop = catch/sum(catch))
  
t %>% 
  group_by(region) %>% 
  # filter(region %notin% c("ISL","GRL")) %>% 

  ggplot(aes(x=year,y=region)) +
  theme_bw() +
  theme(legend.position="right") +
  ggridges::geom_ridgeline(aes(fill=region, height=prop), scale=2) +
  geom_text(data=tt, 
            aes(x=2015, y=region, label=scales::percent(prop, accuracy=1)),
            nudge_y = 0.15) +
  geom_vline(xintercept=2010, linewidth=0.5, linetype="dashed") +
  geom_vline(xintercept=2020, linewidth=0.5, linetype="dashed") +
  labs(x="", y="percentage of catch by Party of total catch", 
       title="relative catch by Party (%) and mean over 2010-2020", fill="Party") +
  # scale_fill_brewer(palette="Set1") +
  scale_fill_brewer(palette="Paired") +
  scale_y_discrete(limits=rev(levels(t$region)))  

# As facet_grid

# t %>% 
#   group_by(region) %>% 
#   mutate(region=factor(region, levels=regions)) %>% 
#   # filter(region %notin% c("ISL","GRL")) %>% 
# 
#   ggplot(aes(x=year,y=prop)) +
#   theme_bw() +
#   theme(legend.position="right") +
#   geom_ribbon(aes(xmin=year, xmax=year,
#                   ymin=0, ymax=prop,
#                   fill=region), colour="black") +
#   geom_text(data=tt, 
#             aes(x=2015, y=0.1, label=scales::percent(prop, accuracy=1))) +
#   geom_vline(xintercept=2010, linewidth=0.2, linetype="dashed") +
#   geom_vline(xintercept=2020, linewidth=0.2, linetype="dashed") +
#   labs(x="", y="perc. of catch", title="relative catch by Party (%) and mean over 2010-2020", fill="Party") +
#   scale_y_continuous(labels=scales::percent) +
#   scale_fill_brewer(palette="Set1") +
#   facet_grid(region ~.)

```

_Figure 2: Mackerel catch proportions of the four parties' in the sharing arrangement._

\newpage

Mackerel catch proportions of all parties' catches

```{r eval=FALSE, fig.align="center", fig.asp=0.8, message=FALSE, warning=FALSE, include=FALSE}

# catch_by_species_year_country_raw %>% distinct(eez) %>% View()

# regions <- c("UK","EU27","NOR","FRO","RUS", "ISL","GRL")
regions <- c("EU27","NOR","UK", "FRO","RUS", "ISL","GRL")

t <-
  catch_by_species_year_country_raw %>% 
  filter(region %in% regions) %>% 
  group_by(species, year, region) %>% 
  summarize(catch = as.integer(sum(catch, na.rm=TRUE))) %>% 
  mutate(prop = catch / sum(catch, na.rm=TRUE)) %>% 
  mutate(region=factor(region, levels=regions))
  
tt <-
  t %>% 
  mutate(region=factor(region, levels=regions)) %>% 
  filter(year %in% 2010:2020) %>% 
  group_by(region) %>% 
  summarise(catch = sum(catch)) %>% 
  ungroup() %>% 
  mutate(prop = catch/sum(catch))
  
t %>% 
  group_by(region) %>% 
  # filter(region %notin% c("ISL","GRL")) %>% 

  ggplot(aes(x=year,y=region)) +
  theme_bw() +
  theme(legend.position="right") +
  ggridges::geom_ridgeline(aes(fill=region, height=prop), scale=2) +
  geom_text(data=tt, 
            aes(x=2015, y=region, label=scales::percent(prop, accuracy=1)),
            nudge_y = 0.15) +
  geom_vline(xintercept=2010, linewidth=0.5, linetype="dashed") +
  geom_vline(xintercept=2020, linewidth=0.5, linetype="dashed") +
  labs(x="", y="percentage of catch by part of total catch", 
       title="relative catch by Party (%) and mean over 2010-2020", fill="Party") +
  scale_fill_brewer(palette="Paired") +
  scale_y_discrete(limits=rev(levels(t$region)))  

```

_Figure 3: Mackerel catch proportions of all parties' catches._

\newpage

Proportion of mackerel catches by party and taken in coastal or international waters

```{r, echo=FALSE, fig.asp=0.6, fig.align="center", message=FALSE, warning=FALSE}

# catch_by_species_year_country_raw %>% distinct(eez) %>% View()

# regions <- c("UK","EU27","NOR","FRO","RUS", "ISL","GRL")
regions <- c("EU27","NOR","UK", "FRO","RUS", "ISL","GRL")

t1 <-
  catch_by_species_year_country_raw_zbyrect %>% 
  filter(region %in% regions) %>% 
  mutate(region=factor(region, levels=regions)) %>% 
  group_by(species, year, region) %>% 
  summarize(catch = as.integer(sum(catch, na.rm=TRUE))) %>% 
  group_by(species, year) %>% 
  mutate(prop = catch / sum(catch, na.rm=TRUE))

t1_tt <-
  t1 %>% 
  mutate(region=factor(region, levels=regions)) %>% 
  filter(year %in% 2010:2020) %>% 
  group_by(region) %>% 
  summarise(catch = sum(catch)) %>% 
  ungroup() %>% 
  mutate(prop = catch/sum(catch))

t2_nest <-
  catch_by_species_year_country_raw_zbyrect %>% 
  filter(region %in% regions) %>% 
  filter(eez == "INT") %>% 
  mutate(region=factor(region, levels=regions)) %>% 
  expand(nesting(species, year), region)  

t2 <-
  catch_by_species_year_country_raw_zbyrect %>% 
  filter(region %in% regions) %>% 
  filter(eez == "INT") %>% 
  mutate(region=factor(region, levels=regions)) %>% 
  
  right_join(t2_nest)  %>% 
  mutate(catch = ifelse(is.na(catch),0,catch)) %>% 
  group_by(species, year, region) %>% 
  summarize(catch_t2 = as.integer(sum(catch, na.rm=TRUE))) %>% 
  left_join(t1, by=c("species","year","region")) %>% 
  mutate(prop_int = catch_t2/catch)

# t2 %>% filter(region=="NOR") %>% View()

t2_tt <-
  bind_rows(
      t2 %>% 
        filter(year %in% 2010:2014) %>% 
        group_by(region) %>% 
        summarise(prop_int = mean(prop_int)) %>% 
        mutate(year=2012),
      t2 %>% 
        filter(year %in% 2015:2020) %>% 
        group_by(region) %>% 
        summarise(prop_int = mean(prop_int)) %>% 
        mutate(year=2018)
  )

p1 <-
  t1 %>% 
  ggplot(aes(x=year,y=region)) +
  theme_bw() +
  theme(legend.position="none") +
  ggridges::geom_ridgeline(aes(fill=region, height=prop), scale=2.5) +
  geom_text(data=t1_tt,
            aes(x=2015, y=region, label=scales::percent(prop, accuracy=1)),
            nudge_y = 0.15) +
  geom_vline(xintercept=2010, linewidth=0.5, linetype="dashed") +
  geom_vline(xintercept=2020, linewidth=0.5, linetype="dashed") +
  labs(x="",y="", title="relative catch by party", fill="Party") +
  scale_fill_brewer(palette="Paired") +
  scale_y_discrete(limits=rev(levels(t1$region)))

# p2 <-
#   t2 %>% 
#   ggplot(aes(x=year,y=region)) +
#   theme_bw() +
#   theme(legend.position="right") +
#   ggridges::geom_ridgeline(aes(fill=region, height=prop_int), scale=1) +
#   geom_text(data=t2_tt,
#             aes(x=year, y=region, label=scales::percent(prop_int, accuracy=1)),
#             nudge_y = 0.15) +
#   geom_vline(xintercept=2010, linewidth=0.5, linetype="dashed") +
#   geom_vline(xintercept=2014.5, linewidth=0.5, linetype="dashed") +
#   geom_vline(xintercept=2020, linewidth=0.5, linetype="dashed") +
#   labs(x="",y="", title="perc. of party's catch taken in international waters", fill="Party") +
#   scale_fill_brewer(palette="Paired") +
#   scale_y_discrete(limits=rev(levels(t1$region)))

p2 <-
  t2 %>%
  # filter(region %notin% c("ISL","GRL")) %>%

  ggplot(aes(x=year,y=prop_int)) +
  theme_bw() +
  theme(legend.position="right") +
  geom_ribbon(aes(xmin=year, xmax=year,
                  ymin=0, ymax=prop_int,
                  fill=region), colour="black") +
  geom_text(data=filter(t2, year==2020),
            aes(x=year-0.1, y=0.1, label=scales::percent(prop_int, accuracy=1)),
            hjust=1) +
  geom_vline(xintercept=2010, linewidth=0.2, linetype="dashed") +
  geom_vline(xintercept=2020, linewidth=0.2, linetype="dashed") +
  labs(x="",y="", title="perc. of party's catch taken in international waters", fill="Party") +
  scale_y_continuous(labels=scales::percent) +
  scale_fill_brewer(palette="Paired") +
  facet_grid(region ~.)

ggpubr::ggarrange(p1,p2, ncol=2, widths = c(1,1,2))


```

_Figure 4: Mackerel catch proportions by party in coastal and international waters._

\newpage

Proportions of catches by month within each year in different economic zones. Only zones where more than contributed more than 1% to the total international catch. Months with less than 0.5% of the annual catch in that zone were removed. 

```{r eval=FALSE, fig.align="center", fig.asp=0.9, message=FALSE, warning=FALSE, include=FALSE}

# catch_by_species_year_country_raw %>% distinct(eez) %>% View()
eezs <- c("EU27","NOR","UK","FRO","RUS", "ISL","GRL","INT")

t <-
  catch_by_species_year_country_raw_zbyrect %>% 
  filter(year >= 2010, year <= 2020) %>% 
  mutate(month=ifelse(ptype=="M", pnum, 1+(pnum-1)*3 )) %>% 
  mutate(eez = ifelse(eez == c("GBR"), "UK", eez) ) %>% 

  group_by(species, eez, month) %>% 
  summarize(catch = as.integer(sum(catch, na.rm=TRUE))) %>% 
  group_by(species, eez) %>% 
  mutate(
    totalcatch = sum(catch, na.rm=TRUE),
    prop       = catch / totalcatch) %>% 
  group_by(species) %>% 
  mutate(allcatch = sum(catch)) %>% 
  filter(totalcatch/allcatch > 0.01)  %>% 
  mutate(eez = factor(eez, levels=eezs))

t %>% 
  ggplot(aes(x=month,y=eez, group=eez)) +
  theme_bw() +
  theme(legend.position="right") +
  ggridges::geom_ridgeline(aes(fill=eez, height=prop), scale=2, rel_min_height = 0.01) +
  # labs(y="relative catch by Coastal State (%)") +
  labs(x="month", y="percentage of catch by month", 
       title="Relative catch by month within zones (2010-2020)", fill="EEZ") +
  scale_fill_brewer(palette="Paired") + 
  scale_x_continuous(breaks=seq(1,12,1)) +
  scale_y_discrete(limits=rev(levels(t$eez)))


```

_Figure 5: Proportions of catches by month within each year in different economic zones._

\newpage

Mackerel catch proportions of of the EU27 fleet

```{r eval=FALSE, fig.align="center", fig.asp=0.8, message=FALSE, warning=FALSE, include=FALSE}

# catch_by_species_year_country_raw %>% distinct(eez) %>% View()

t <-
  catch_by_species_year_country_raw %>% 
  filter(region %in% "EU27") %>% 
  filter(year >= 2003) %>% 
  mutate(month=ifelse(ptype=="M", pnum, 1+(pnum-1)*3 )) %>% 
  group_by(species, year, month) %>% 
  summarize(catch = as.integer(sum(catch, na.rm=TRUE))) %>% 
  mutate(prop = catch / sum(catch, na.rm=TRUE))

  
t %>% 
  ggplot(aes(x=month,y=year, group=year)) +
  theme_bw() +
  theme(legend.position="right") +
  ggridges::geom_ridgeline(aes(height=prop), 
                           scale=2.5, fill=RColorBrewer::brewer.pal(8, "Paired")[1]) +
  labs(x="", y="proportion of catch by month", 
       title="relative catch by month of the EU27 fleet") +
  scale_fill_brewer(palette="Set1") +
  scale_x_continuous(breaks=seq(1,12,1)) +
  scale_y_reverse() 

```

_Figure x: Mackerel catch proportions of all parties' catches._

Mackerel catch proportions of of the EU27 fleet

```{r, echo=FALSE, fig.asp=0.8, fig.align="center", message=FALSE, warning=FALSE}

# catch_by_species_year_country_raw %>% distinct(eez) %>% View()

regions <- c("EU27","NOR","UK", "FRO","RUS", "ISL","GRL")

t <-
  catch_by_species_year_country_raw %>% 
  # filter(region %in% "EU27") %>% 
  filter(year >= 2003) %>% 
  mutate(month=ifelse(ptype=="M", pnum, 1+(pnum-1)*3 )) %>% 
  mutate(region = factor(region, levels=regions)) %>% 
  group_by(species, region, year, month) %>% 
  summarize(catch = as.integer(sum(catch, na.rm=TRUE))) %>% 
  group_by(species, region, year) %>% 
  mutate(prop = catch / sum(catch, na.rm=TRUE)) %>% 
  mutate(prop = ifelse(prop < 0.005, NA, prop)) 


  
t %>% 
  ggplot(aes(x=month,y=year, group=year)) +
  theme_bw() +
  theme(legend.position="right") +
  ggridges::geom_ridgeline(aes(height=prop, fill=region), 
                           scale=2.5) +
  labs(x="month", y="proportion of catch by month", 
       title="Relative catch by month and year by different parties", fill="Party") +
  scale_fill_brewer(palette="Paired") +
  scale_y_reverse() +
  scale_x_continuous(breaks=seq(1,12,1)) +
  facet_wrap(~region, ncol=4)

```

_Figure x: Mackerel catch proportions of all parties' catches._


\newpage

Proportions of catches by month within each year in different economic zones. Only zones where more than contributed more than 1% to the total international catch. Months with less than 0.5% of the annual catch in that zone were removed. 

```{r, echo=FALSE, fig.asp=0.9, fig.align="center", message=FALSE, warning=FALSE}

# catch_by_species_year_country_raw %>% distinct(eez) %>% View()
eezs <- c("EU27","NOR","UK","FRO","RUS", "ISL","GRL","INT")

t <-
  catch_by_species_year_country_raw_zbyrect %>% 
  filter(year >= 2000) %>% 
  mutate(month=ifelse(ptype=="M", pnum, 1+(pnum-1)*3 )) %>% 
  mutate( eez = ifelse(eez %in% c("GBR"), "UK", eez) ) %>% 

  group_by(species, eez, year, month) %>% 
  summarize(catch = as.integer(sum(catch, na.rm=TRUE))) %>% 
  group_by(species, eez, year) %>% 
  mutate(prop = catch / sum(catch, na.rm=TRUE)) %>% 
  group_by(species, eez) %>% 
  mutate(eezcatch = sum(catch)) %>% 
  group_by(species) %>% 
  mutate(totalcatch = sum(catch)) %>% 
  filter(eezcatch/totalcatch > 0.01) %>% 
  mutate(prop = ifelse(prop < 0.005, NA, prop)) %>% 
  mutate(eez = factor(eez, levels=eezs))

t %>% 
  ggplot(aes(x=month,y=year, group=year)) +
  theme_bw() +
  theme(legend.position="right") +
  ggridges::geom_ridgeline(aes(fill=eez, height=prop), scale=2, rel_min_height = 0.01) +
  labs(x="month", y="proportion of catch by month", 
       title="Relative catch by month and year within different zones", fill="EEZ") +
  scale_fill_brewer(palette="Paired") +
  scale_y_reverse() +
  scale_x_continuous(breaks=seq(1,12,1)) +
  facet_wrap(~eez, ncol=4)

```

_Figure 5: Proportions of catches by month within each year in different economic zones._

\newpage


\newpage

Proportions of catches by month within each year within the EU27 EEZ, separated into North Sea, southern and western area. Months with less than 0.5% of the annual catch in that zone were removed. 

```{r, echo=FALSE, fig.asp=0.9, fig.align="center", message=FALSE, warning=FALSE}

t <-
  catch_by_species_year_country_raw_zbyrect %>% 
  filter(year >= 2000) %>% 
  mutate(month=ifelse(ptype=="M", pnum, 1+(pnum-1)*3 )) %>% 
  filter(eez %in% c("EU27")) %>% 
  mutate(division   = 
           geo_inside(lon=lon, 
                      lat=lat, 
                      map=fao[fao@data$F_LEVEL=="DIVISION",], 
                      variable="F_DIVISION")) %>%
  mutate(area = case_when(
    division %in% c("27.8.c","27.9.a", "27.9.b") ~ "south",
    division %in% c("27.4.c","27.4.b")           ~ "northsea",
    TRUE                                         ~ "west"
  )) %>% 
  mutate(area = factor(area, levels=c("south","west","northsea"))) %>% 
  group_by(species, area, year, month) %>% 
  summarize(catch = as.integer(sum(catch, na.rm=TRUE))) %>% 
  group_by(species, area, year) %>% 
  mutate(prop = catch / sum(catch, na.rm=TRUE)) %>% 
  mutate(prop = ifelse(prop < 0.005, NA, prop)) 

t %>% 
  ggplot(aes(x=month,y=year, group=year)) +
  theme_bw() +
  theme(legend.position="right") +
  ggridges::geom_ridgeline(aes(fill=area, height=prop), scale=2, rel_min_height = 0.01) +
  # labs(title="relative catch by month and area within EU27 EEZ") +
  labs(x="", y="proportion of catch by month", 
       title="relative catch by month and area within the EU27 EEZ", fill="area") +
  scale_fill_brewer(palette="Paired") +
  scale_y_reverse() +
  scale_x_continuous(breaks=seq(1,12,1)) +
  # scale_alpha_continuous(fill=alpha) +
  facet_wrap(~area)

```

_Figure 6: Proportions of catches by month within each year within the EU27 EEZ, separated into North Sea, southern and western area._

