---
output: 
  word_document:
    reference_docx: ../report_template_v1.5.dotx
---

```{r setup, include=FALSE}

# =======================================================================================
# WGWIDE Stock summaries.Rmd
# 
# Generate stock summary report for WGWIDE
#
# 03/09/2017 first version
# 06/08/2018 UPdated and prepared for WGWIDE 2018
# 07/09/2018 Added link to SAG database for plotting. Currently reading from downloaded data. 
#            Needs to be made through live connection. 
# =======================================================================================

require("knitr")
knitr::opts_chunk$set(echo = FALSE,	message = FALSE,	warning = FALSE,	comment = "",	crop = TRUE )
knitr::opts_chunk$set(fig.width=10) 

rm(list=ls())

library(tidyverse)    # tidying packages
library(readxl)       # read excel
library(lubridate)
library(sf)
library(sp)
library(RColorBrewer)
library(pander)

library(cowplot)                   # for plotting table as image (bio data)

# library(gisland)      # devtools::install_github("einarhjorleifsson/gisland", dependencies = TRUE)

library(maps)
# library(mapdata)
# library(mapproj)
library(lubridate)
library(viridis)

# library(animation)
# library(gganimate)

# source my utils
source("../../prf/R/my utils.r")
source("../../gisland/r/geo_inside.R")

# Data path
datapath <- "C:/Users/Martin/Onedrive - PFA/Documents/iWGWIDE/2022/06. Data/_catch_by_rectangle/"
# datapath <- "C:/DATA/Onedrive - PFA/Documents/iWGWIDE/iWGWIDE 2021/06. Data/_catch_by_rectangle/"
# datapath <- "//community.ices.dk@SSL/DavWWWRoot/ExpertGroups/WGWIDE/2019 Meeting docs/06. Data/_catch_by_rectangle"

onedrive <- get_onedrive()
load(file.path(onedrive, "rdata/eez.RData"))
load(file.path(onedrive, "rdata/fao.RData"))

load(file.path(onedrive, "rdata/eez.df.RData"))
load(file.path(onedrive,"rdata/icesrectangles.df.RData"))
icesrect <- 
  icesrectangles.df %>% 
  group_by(rect) %>% 
  distinct(rect, .keep_all = TRUE) %>% 
  mutate(
    lon  = (WEST + EAST)/2,
    lat  = (SOUTH + NORTH)/2
  )
# icesrect <-
#   icesrectangles.df %>% 
#   distinct(rect, lat=SOUTH, lon=WEST) %>% 
#   dplyr::select(rect, lon, lat)
# write.csv(icesrect, file="icesrect.csv", row.names = FALSE)
# icesrect <- read.csv(file=file.path(datapath, "icesrect.csv")) %>% mutate(across(c("lat","lon"), as.numeric))

# list the available files within the directory
files.list <- list.files(path=datapath, 
                         pattern="WGWIDE.*xlsx",
                         full.names=TRUE )

# read the files
f <- files.list[1]

catch_by_species_year_country_raw <- data.frame(stringsAsFactors = FALSE)

for (f in files.list) {
  print(f)
  catch_by_species_year_country_raw <-
    bind_rows(
      catch_by_species_year_country_raw,
      read_excel(f, col_names=TRUE, col_types="text") %>% 
        lowcase() %>% 
        dplyr::mutate(across(c("catch", "lat","lon"), as.numeric ))  %>%
        dplyr::mutate(across(c("year"), as.integer)) %>% 
        dplyr::mutate(pnum = as.integer(gsub("quarter","", pnum))) %>% 
        dplyr::mutate(ptype = toupper(ptype)) %>% 
        
        # change month to quarter
        dplyr::mutate(pnum  = ifelse(ptype=="M", ceiling(as.numeric(pnum)/3), pnum)) %>% 
        dplyr::mutate(ptype = ifelse(ptype=="M", "Q", ptype)) %>% 
        
        filter(catch >0) 
    )  
}


catch_by_species_year_country_raw <-
  catch_by_species_year_country_raw %>% 
  
  # replace lat long with values from icesrect (bottom-left points of rectangles)
  dplyr::select(-lat, -lon) %>% 
  dplyr::left_join(icesrect, by=c("rect")) %>% 
  dplyr::filter((species == "hom" & lon > -25) | (species != "hom")) %>% 
  dplyr::mutate(eez        = geo_inside(lon=lon, lat=lat, map=eez, variable="Territory1")) %>%
  dplyr::mutate(division   = geo_inside(lon=lon, lat=lat, map=fao[fao@data$F_LEVEL=="DIVISION",], variable="F_DIVISION")) %>%
  
  # replace country names
  dplyr::mutate(country = ifelse(country %in% c("UKE","GBR.E"), "GBR.EW", country)) %>% 
  dplyr::mutate(country = ifelse(country %in% c("BES","SPA")    , "ESP", country)) %>% 
  dplyr::mutate(country = ifelse(country %in% c("POR")          , "PRT", country)) %>% 
  dplyr::mutate(country = ifelse(country %in% c("GER")          , "DEU", country))  

# su(catch_by_species_year_country_raw$country)
# catch_by_species_year_country_raw %>% filter(year==2021) %>% distinct(species)

# mapplots::ices.rect(catch_by_species_year_country_raw$rect) %>% View()

# catch by year
catch_by_species_year <- 
  catch_by_species_year_country_raw %>% 
  group_by(species, year, rect, lat, lon) %>% 
  summarize(catch = sum(catch, na.rm=TRUE)) %>% 
  
  mutate(catch_interval = cut(as.integer(catch), breaks=c(1 %o% 10^(0:12)), dig.lab=10 ) ) %>% 
  mutate(catch_interval = gsub(" ","", catch_interval)) %>% 
  filter(!is.na(catch_interval)) %>% 
  filter(grepl("\\,", catch_interval)) %>%   
  group_by(species, year, rect, lat, lon, catch_interval) %>% 
  summarize(catch = sum(catch, na.rm=TRUE))

   # map_data("worldHires", xlim = xlim, ylim = ylim) %>%

catch_by_species_year_quarter_division <- 
  catch_by_species_year_country_raw %>% 
  filter(ptype == "Q") %>% 
  rename(quarter = pnum) %>% 
  mutate(stock = as.character(NA)) %>% 
  mutate(stock = case_when(
    species == "HER"     ~ "AS herring", 
    species == "MAC"     ~ "NEA mackerel",
    species == "WHB"     ~ "NEA blue whiting",
    species == "HOM" & division %in% c("27.9.a")                    ~ "Southern horse mackerel",
    species == "HOM" & division %in% c("27.4.b","27.4.c","27.7.d")  ~ "Northsea horse mackerel",
    species == "HOM" & division %in% c("27.4.a") & quarter %in% 1:2 ~ "Northsea horse mackerel",
    species == "HOM" & is.na(stock)                                 ~ "Western horse mackerel")
  ) %>%  

  group_by(species, stock, year, quarter, division, rect, lat, lon) %>% 
  summarize(catch = sum(catch, na.rm=TRUE)) %>% 
  
  mutate(catch_interval = cut(as.integer(catch), breaks=c(1 %o% 10^(0:12)), dig.lab=10 ) ) %>% 
  mutate(catch_interval = gsub(" ","", catch_interval)) %>% 
  filter(!is.na(catch_interval)) %>% 
  filter(grepl("\\,", catch_interval)) %>%   
  group_by(species, stock, year, quarter, division, rect, lat, lon, catch_interval) %>% 
  summarize(catch = sum(catch, na.rm=TRUE))

# catch_by_species_year_quarter_division %>% ungroup() %>% distinct(stock) %>% View()

# ================================================================================
# FUNCTIONS
# ================================================================================

table_catch_by_year_country <- function(myspecies="MAC", myyears=NA) {
  
  # t <-
    catch_by_species_year_country_raw %>% 
    filter(species == myspecies) %>% 
    {if ((length(myyears) == 1) & all(is.na(myyears))) {
      filter(.)
    } else {
      filter(., year %in% myyears)
    }} %>% 

    group_by(species, year, country) %>% 
    summarize(catch = as.integer(sum(catch, na.rm=TRUE))) %>% 
    reshape2::dcast(country ~ year, value.var="catch", sum, margins=c("year","country")) %>% 
    # { if (ncol(.) > split) }
    pander::pandoc.table(.,
               style = "simple",
               split.tables=120, 
               justify = "right",
               missing=".")

} # end of table_catch_by_year_country

# table_catch_by_year_country (myspecies="MAC")

table_catch_by_year_species <- function(myyears=NA) {
  
  catch_by_species_year_country_raw %>% 
    {if ((length(myyears) == 1) & all(is.na(myyears))) {
      filter(.)
    } else {
      filter(., year %in% myyears)
    }} %>% 

    group_by(species, year) %>% 
    summarize(catch = as.integer(sum(catch, na.rm=TRUE))) %>% 
    reshape2::dcast(species ~ year, value.var="catch", sum, margins=c("year","country")) %>% 
    # { if (ncol(.) > split) }
    pander::pandoc.table(.,
               style = "simple",
               split.tables=120, 
               justify = "right",
               missing=".")

} # end of table_catch_by_year_species

# table_catch_by_year_species()


# myspecies = "HOM"; myyears = 1998:2021; ncol=6
# myspecies = "MAC"; myyears = 1998
# myspecies = "MAC"; myyears = NA
# myspecies = "HOM"; myyears = NA

plot_catch_by_year <- function(myspecies="MAC", myyears=NA, plot_catch=TRUE, plot_ssb=FALSE, ncol=6, 
                               xlim=NA, ylim=NA, add_midpoint=FALSE) {
  
  catch2 <-
    catch_by_species_year %>% 
    drop_na(lat, lon) %>% 
    filter(species == myspecies) %>% 
    {if ((length(myyears) == 1) & all(is.na(myyears))) {
      filter(.)
    } else {
      filter(., year %in% myyears)
    }} %>% 
    
    {if (!all(is.na(myyears))) bind_rows(., data.frame(species=myspecies, year = myyears, catch_interval="(1,10]", catch=0)) else (.) }

  if(all(is.na(xlim))) {xlim <- range(catch2$lon, na.rm=TRUE)}
  if(all(is.na(ylim))) {ylim <- range(catch2$lat, na.rm=TRUE)}

  mp <- catch2 %>% 
    group_by(year) %>% 
    summarise(
      lon = weighted.mean(lon, catch),
      lat = weighted.mean(lat, catch)
    )

  # print(mp)  

  tc <-
    catch2 %>% 
    group_by(species, year) %>% 
    summarize(catch = sum(catch, na.rm=TRUE)) %>% 
    mutate(catch  = as.integer(catch) ) %>% 
    
    group_by(species) %>% 
    mutate(catch2 = ylim[1] + (catch / max(catch, na.rm=TRUE) * (ylim[2] - ylim[1])) )
  
  ggplot2::map_data("world", xlim = xlim, ylim = ylim) %>% 
    ggplot(aes(long, lat)) +
    geom_polygon(aes(long, lat, group = group), fill = "gray") +
    coord_quickmap(xlim = xlim, ylim = ylim, expand = FALSE) +
    
    theme_publication() +
    theme(panel.border     = element_rect(colour="black" , size=0.2),
          panel.grid.major = element_blank(),
          strip.background = element_rect(colour="black", size =0.2),
          plot.margin      = unit(c(0,0,0,0),"cm"),
          plot.title       = element_text(hjust=0, vjust=0, size=10),
          axis.text        = element_text(size=6),
          legend.key.width = unit(0.4, "cm"), 
          axis.title       = element_blank()) +
    
    geom_tile(data=catch2, aes(x=lon, y=lat, fill = catch_interval, height=0.5, width=1), colour=NA, alpha=1.0) +
    scale_fill_viridis(option = "plasma", direction = -1, discrete=TRUE) +

    {if(add_midpoint==TRUE) geom_point(data=mp, aes(x=lon, y=lat), colour="black", size=2, inherit.aes = FALSE)} +
    {if(add_midpoint==TRUE) geom_hline(data=mp, aes(yintercept=lat), colour="black", inherit.aes = FALSE)} +
    {if(add_midpoint==TRUE) geom_vline(data=mp, aes(xintercept=lon), colour="black", inherit.aes = FALSE)} +
    
    geom_label(data=tc, aes(label=paste0("C ",catch)), x=-Inf, y=Inf, hjust=0, vjust=1, inherit.aes=FALSE) +
  
    
    labs(x = NULL, y = NULL, size = "tons", title=unique(catch2$species)) +
    guides(fill = guide_legend(nrow = 1, title="catch (tonnes)")) +
    facet_wrap( ~ year, ncol=ncol)

} # end of plot_catch_by_year

# myspecies = "HOM"; myyears = 2015:2021; mydivision="27.4.a";plot_catch=TRUE; xlim=NA; ylim=NA; add_midpoint=FALSE; plot_eez=TRUE

plot_catch_by_division_year_quarter <- 
  function(myspecies="MAC", myyears=NA, mydivision=c("27.4.a"), plot_eez=FALSE, plot_catch=TRUE, 
           xlim=NA, ylim=NA, add_midpoint=FALSE) {
  
  catch2 <-
    catch_by_species_year_quarter_division %>% 
    drop_na(division, quarter) %>% 
    filter(catch > 0) %>% 
    filter(species == myspecies) %>% 
    filter(division %in% mydivision) %>% 
    {if ((length(myyears) == 1) & all(is.na(myyears))) {
      filter(.)
    } else {
      filter(., year %in% myyears)
    }}  
    
    # {if (!all(is.na(myyears))) bind_rows(., data.frame(species=myspecies, year = myyears, catch_interval="(1,10]", catch=0)) else (.) }

  if(all(is.na(xlim))) {xlim <- range(catch2$lon, na.rm=TRUE)}
  if(all(is.na(ylim))) {ylim <- range(catch2$lat, na.rm=TRUE)}

  mp <- catch2 %>% 
    group_by(year, quarter, division) %>% 
    summarise(
      lon = weighted.mean(lon, catch),
      lat = weighted.mean(lat, catch)
    )

  # print(mp)  

  tc <-
    catch2 %>% 
    group_by(species, year, quarter, division) %>% 
    summarize(catch = sum(catch, na.rm=TRUE)) %>% 
    mutate(catch  = as.integer(catch) ) %>% 
    
    group_by(species) %>% 
    mutate(catch2 = ylim[1] + (catch / max(catch, na.rm=TRUE) * (ylim[2] - ylim[1])) )
  
  ggplot2::map_data("world", xlim = xlim, ylim = ylim) %>% 
    ggplot(aes(long, lat)) +
    geom_polygon(aes(long, lat, group = group), fill = "gray") +
    coord_quickmap(xlim = xlim, ylim = ylim, expand = FALSE) +
    
    theme_publication() +
    theme(panel.border     = element_rect(colour="black" , size=0.2),
          panel.grid.major = element_blank(),
          strip.background = element_rect(colour="black", size =0.2),
          plot.margin      = unit(c(0,0,0,0),"cm"),
          plot.title       = element_text(hjust=0, vjust=0, size=10),
          axis.text        = element_text(size=6),
          legend.key.width = unit(0.4, "cm"), 
          axis.title       = element_blank()) +
    
    geom_tile(data=catch2, aes(x=lon, y=lat, fill = catch_interval, height=0.5, width=1), colour=NA, alpha=1.0) +
    scale_fill_viridis(option = "plasma", direction = -1, discrete=TRUE) +
    
    {if(plot_eez==TRUE) geom_polygon(data=eez.df, aes(long, lat, group = group), fill = NA, colour="gray")} +

    {if(add_midpoint==TRUE) geom_point(data=mp, aes(x=lon, y=lat), colour="black", size=2, inherit.aes = FALSE)} +
    {if(add_midpoint==TRUE) geom_hline(data=mp, aes(yintercept=lat), colour="black", inherit.aes = FALSE)} +
    {if(add_midpoint==TRUE) geom_vline(data=mp, aes(xintercept=lon), colour="black", inherit.aes = FALSE)} +
    
    geom_label(data=tc, aes(label=paste0("C ",catch)), x=-Inf, y=Inf, hjust=0, vjust=1, inherit.aes=FALSE) +
  
    
    labs(x = NULL, y = NULL, size = "tons", title=unique(catch2$species)) +
    guides(fill = guide_legend(nrow = 1, title="catch (tonnes)")) +
    facet_grid(quarter ~ year)

} # end of plot_catch_by_division_year_quarter



plot_catch_by_year_and_country <- function() {
  
} # end of plot_catch_by_year_and_country

plot_catch_by_year_animated <- function() {
  
} # end of plot_catch_by_year_animated

# myspecies="HER"; myyears = 2003:2021
plot_lat_long <- function(myspecies="MAC", myyears=NA) {
  
  catch2 <-
    catch_by_species_year %>% 
    drop_na(lat, lon) %>% 
    filter(species == myspecies) %>% 
    {if ((length(myyears) == 1) & all(is.na(myyears))) {
      filter(.)
    } else {
      filter(., year %in% myyears)
    }} %>% 
    
    {if (!all(is.na(myyears))) 
      bind_rows(., data.frame(species=myspecies, year = myyears, catch_interval="(1,10]", catch=0)) else (.)} %>% 
    
    group_by(species, year) %>% 
    summarise(
      lon = weighted.mean(lon, catch),
      lat = weighted.mean(lat, catch)) 

  # View(catch2)
  
  p1 <-
    catch2 %>% 
    ggplot(aes(year, lat)) +
    theme_publication() +

    geom_line() +
    geom_point() +
    
    labs(x = "year", y = "lat", size = "tons", title = paste(myspecies, "mean latitude")) 
  
  p2 <-
    catch2 %>% 
    ggplot(aes(year, lon)) +
    theme_publication() +

    geom_line() +
    geom_point() +
    coord_flip() +
    
    labs(x = "year", y = "lon", size = "tons", title=paste(myspecies, "mean longitude")) 

  cowplot::plot_grid(p1, p2, ncol=2, scale=0.98, align="hv",
                 rel_widths = c(1.0, 1.0), rel_heights = c(1.0, 1.0))

  } # end of plot_catch_by_year


```
Working document 10, WGWIDE 2021

**Full time-series of catch by rectangle**

Martin Pastoors, `r format(Sys.time(), '%d/%m/%Y')`

**Introduction**

WGWIDE and its precursors WGMHSA and WGNPBW have been publishing catch per rectangle plots in their reports for many years already. Catch by rectangle has been compiled by WG members and generally provide a WG estimate of catch per rectangle. In most cases the information is available by quarter whereas most recently, the data has been requested by month. Previously, the catch by rectangle has mostly presented for one single year in the WG reports. Here, we collated all the catch by rectangle data that is available for herring, blue whiting, mackerel and horse mackerel for as many years as available. 

**Results**

An overview of the available catches by species and year is shown in the text table below. For horse mackerel and mackerel, a long time series is available, starting in 2001 (HOM) and 1998 (MAC). The time series for herring and blue whiting are shorter (starting in 2011) although additional information could be derived from earlier WG reports. 

```{r, echo=FALSE, fig.asp=1.4, fig.align="center", message=FALSE, warning=FALSE}

table_catch_by_year_species()

```

For each species an overview table is presented of catch by country and year and a figure with catch by rectangle and year. Catches by rectangle have been grouped in logarithmic classes (1-10, 10-100 etc). 

**Discussion**

While the aggregation and presentation of the catch per rectangle data for mackerel, horse mackerel, blue whiting and atlanto-scandian herring does not constitute rocket-science, it does provide us with meaningful insights into the changes of catching areas over time. This could be relevant also in understanding the impacts of climate change on fisheries and in relating changes in the distribution of prey or predator species (e.g. bluefin tuna). As such, these graphical representations of catching areas provide a useful addition to the WG report. 

One important check that still needs to be carried out is the check on data availability by country and year that may not be consistent over the time series. Making the time-series complete would improve the useability of the information. 

\newpage

**Mackerel**

```{r, echo=FALSE, fig.asp=1.4, fig.align="center", message=FALSE, warning=FALSE}

table_catch_by_year_country(myspecies="MAC")

```

_Table 1: Catch of mackerel (tonnes) included in the rectangle data by year and country_

```{r, echo=FALSE, fig.asp=1.2, fig.align="center", message=FALSE, warning=FALSE}

plot_catch_by_year(myspecies="MAC", myyears = 2003:2021, ncol=5, add_midpoint = TRUE)

```

_Figure 1: Catch of mackerel (tonnes) by year and rectangle_

\newpage

```{r, echo=FALSE, fig.asp=0.6, fig.align="center", message=FALSE, warning=FALSE}

plot_lat_long(myspecies="MAC", myyears = 2003:2021)

```

\newpage

**Horse Mackerel**

```{r, echo=FALSE, fig.asp=1.4, fig.align="center", message=FALSE, warning=FALSE}

table_catch_by_year_country(myspecies="HOM")

```

_Table 2: Catch of horse mackerel (tonnes) included in the rectangle data by year and country_

```{r, echo=FALSE, fig.asp=1.2, fig.align="center", message=FALSE, warning=FALSE}

plot_catch_by_year(myspecies="HOM", myyears = 2003:2021, ncol=5, xlim=c(-25,10))


# plot_catch_by_division_year_quarter (myspecies = "HOM", myyears = 2015:2021, mydivision="27.4.a",plot_catch=TRUE, xlim=NA, ylim=NA, add_midpoint=FALSE, plot_eez=TRUE) 

```

_Figure 2: Catch of horse mackerel (tonnes) by year and rectangle_

\newpage

```{r echo=FALSE, fig.align="center", fig.asp=0.6, message=FALSE, warning=FALSE}

myyears <- 2016:2021

catch2 <-
  catch_by_species_year_quarter_division %>% 
  drop_na(division, quarter) %>% 
  filter(catch > 0) %>% 
  filter(species == "HOM") %>%
  {if ((length(myyears) == 1) & all(is.na(myyears))) {
    filter(.)
  } else {
    filter(., year %in% myyears)
  }}  

  # mystock <- "Western horse mackerel"
  i <- 0
  for (mystock in c( "Northsea horse mackerel", "Western horse mackerel", "Southern horse mackerel")) {
    i <- i + 1
    
    xlim <- range(filter(catch2, stock==mystock)$lon, na.rm=TRUE)
    ylim <- range(filter(catch2, stock==mystock)$lat, na.rm=TRUE)
    
    
    mp <- 
      catch2 %>% 
      filter(stock==mystock) %>% 
      group_by(year, stock) %>% 
      summarise(
        lon = weighted.mean(lon, catch),
        lat = weighted.mean(lat, catch)
      )
    
    # print(mp)  
  
    tc <-
      catch2 %>% 
      filter(stock==mystock) %>% 
      group_by(species, stock, year) %>% 
      summarize(catch = sum(catch, na.rm=TRUE)) %>% 
      mutate(catch  = as.integer(catch) ) %>% 
      
      group_by(species, stock) %>% 
      mutate(catch2 = ylim[1] + (catch / max(catch, na.rm=TRUE) * (ylim[2] - ylim[1])) )
    
    print (
      ggplot2::map_data("world", xlim = xlim, ylim = ylim) %>% 
      ggplot(aes(long, lat)) +
      geom_polygon(aes(long, lat, group = group), fill = "gray") +
      coord_quickmap(xlim = xlim, ylim = ylim, expand = FALSE) +
      
      theme_publication() +
      theme(panel.border     = element_rect(colour="black" , size=0.2),
            panel.grid.major = element_blank(),
            strip.background = element_rect(colour="black", size =0.2),
            plot.margin      = unit(c(0,0,0,0),"cm"),
            plot.title       = element_text(hjust=0, vjust=0, size=10),
            axis.text        = element_text(size=6),
            legend.key.width = unit(0.4, "cm"), 
            axis.title       = element_blank()) +
      
      geom_tile(data=filter(catch2, stock == mystock), 
                aes(x=lon, y=lat, fill = catch_interval, height=0.5, width=1), colour=NA, alpha=1.0) +
      scale_fill_viridis(option = "plasma", direction = -1, discrete=TRUE) +
      
      geom_polygon(data=eez.df, aes(long, lat, group = group), fill = NA, colour="gray") +
  
      geom_point(data=mp, aes(x=lon, y=lat), colour="black", size=2, inherit.aes = FALSE) +
      geom_hline(data=mp, aes(yintercept=lat), colour="black", inherit.aes = FALSE) +
      geom_vline(data=mp, aes(xintercept=lon), colour="black", inherit.aes = FALSE) +
      
      geom_label(data=tc, aes(label=paste0("C ",catch)), x=-Inf, y=Inf, hjust=0, vjust=1, inherit.aes=FALSE) +
    
      labs(x = NULL, y = NULL, size = "tons", title=paste(unique(tc$species), unique(tc$stock))) +
      guides(fill = guide_legend(nrow = 1, title="catch (tonnes)")) +
      facet_grid(stock ~ year)
    )
  }


# catch_by_species_eez_year %>% filter(is.na(eez)) %>% View()

```

\newpage

**Blue whiting**

```{r, echo=FALSE, fig.asp=1.4, fig.align="center", message=FALSE, warning=FALSE}

table_catch_by_year_country(myspecies="WHB")

```

_Table 3: Catch of blue whiting (tonnes) included in the rectangle data by year and country_

```{r, echo=FALSE, fig.asp=0.9, fig.align="center", message=FALSE, warning=FALSE}

plot_catch_by_year(myspecies="WHB", myyears = 2011:2021, ncol=4, add_midpoint = TRUE)

```

_Figure 3: Catch of blue whiting (tonnes) by year and rectangle_

\newpage

```{r, echo=FALSE, fig.asp=0.6, fig.align="center", message=FALSE, warning=FALSE}

plot_lat_long(myspecies="WHB", myyears = 2003:2021)

```


\newpage

**Atlanto-scandian herring**

```{r, echo=FALSE, fig.asp=1.4, fig.align="center", message=FALSE, warning=FALSE}

table_catch_by_year_country(myspecies="HER")

```

_Table 4: Catch of Atlanto-scandian herring (tonnes) included in the rectangle data by year and country_

```{r, echo=FALSE, fig.asp=0.8, fig.align="center", message=FALSE, warning=FALSE}

plot_catch_by_year(myspecies="HER", myyears = 2011:2021, ncol=4)

```

_Figure 4: Catch of Atlanto-scandian herring (tonnes) by year and rectangle_

\newpage

```{r, echo=FALSE, fig.asp=0.6, fig.align="center", message=FALSE, warning=FALSE}

plot_lat_long(myspecies="HER", myyears = 2003:2021)

```


```{r, echo=FALSE, fig.asp=0.8, fig.align="center", message=FALSE, warning=FALSE}

# catch by year
# catch_by_species_eez_year <- 
#   catch_by_species_year_country_raw %>% 
#   group_by(species, year, eez, rect, lat, lon) %>% 
#   summarize(catch = sum(catch, na.rm=TRUE)) 


```



