---
output: 
  word_document:
    reference_docx: ../report_template_v1.5.dotx
---

```{r setup, include=FALSE}

# =======================================================================================
# Catch by rectangle by species overview.Rmd
# 
# Generate stock summary report for WGWIDE
#
# 03/09/2017 first version
# 06/08/2018 UPdated and prepared for WGWIDE 2018
# 07/09/2018 Added link to SAG database for plotting. Currently reading from downloaded data. 
#            Needs to be made through live connection. 
# 16/11/2021 Adapted for Coastal States meeting
# 04/02/2021 Adapted from blue whiting code; made more generic
# =======================================================================================

# Find package dependencies in the codes in this folder
# d <- renv::dependencies()

require("knitr")
knitr::opts_chunk$set(echo = FALSE,	message = FALSE,	warning = FALSE,	comment = "",	crop = TRUE )
knitr::opts_chunk$set(fig.width=10) 

rm(list=ls())

library(tidyverse)    # tidying packages
library(readxl)       # read excel
library(lubridate)
library(sf)
library(sp)
library(RColorBrewer)
library(pander)

# library(gisland)      # devtools::install_github("einarhjorleifsson/gisland", dependencies = TRUE)

library(maps)
# library(mapdata)
# library(mapproj)
library(lubridate)
library(viridis)

# library(animation)
# library(gganimate)

# source my utils
source("../../prf/R/my utils.r")
source("../../gisland/r/geo_inside.R")

dropboxdir <- paste(get_dropbox(), "/iAdvice", sep="")

# Load dataset
load(file=paste(dropboxdir, "/rdata/iAdvice.RData",sep=""))

# Data path
datapath <- "C:/DATA/Onedrive - PFA/Documents/iWGWIDE/2021/06. Data/"

onedrive <- get_onedrive()

load(file.path(onedrive, "rdata/world.df.RData"))
load(file.path(onedrive, "rdata/eez.df.RData"))
load(file.path(onedrive, "rdata/fao.df.RData"))
load(file.path(onedrive, "rdata/depth200.df.RData"))
load(file.path(onedrive, "rdata/icesrectangles.df.RData"))

load(file.path(onedrive, "rdata/eez.RData"))

load(file=file.path(onedrive,"rdata/iho.eez.na.sf.RData"))

# ices rectangles
icesrect <- 
  icesrectangles.df %>% 
  group_by(rect) %>% 
  distinct(rect, .keep_all = TRUE) %>%
  dplyr::select(-long) %>% 
  mutate(
    lon  = (WEST + EAST)/2,
    lat  = (SOUTH + NORTH)/2
  )
# icesrect <- read.csv(file=file.path(datapath, "icesrect.csv")) %>% mutate(across(c("lat","lon"), as.numeric))

# Zbycountry: proportion of zones per rectangle
zbycountry <- 
  readxl::read_xlsx(path=file.path(datapath, "_catch_by_rectangle/ZbyCountry_withUK.xlsx"),
                    col_types="text",
                    .name_repair = "unique") %>% 
  lowcase() %>% 
  mutate(across(c("lat","lon","area", "prop"), as.numeric)) %>% 
  dplyr::select(-geometry)
# zbycountry %>% distinct(zone) %>% View()

zbycoastalstate <- 
  readxl::read_xlsx(path=file.path(datapath, "_catch_by_rectangle/ZbyCoastalstate_withUK.xlsx"),
                    col_types="text",
                    .name_repair = "unique") %>% 
  lowcase() %>% 
  mutate(across(c("lat","lon","area", "prop"), as.numeric)) %>% 
  dplyr::select(-geometry)
# zbycoastalstate %>% distinct(zone) %>% View()

#  Selections ==============================================================================================

# myspecies     <- "WHB"
# mystock       <- "whb-comb"
# myspeciesname <- "Blue whiting"
# myfile        <- file.path(datapath, "_catch_by_rectangle/WGWIDE catchesbyrect WHB.xlsx") 
# icescatchfile <- file.path(datapath, "whb.27.1-91214/ICES WHB catch by country.xlsx")
# acdrfile      <- "C:/TEMP/WHB compare datasets2.xlsx"

myspecies     <- "MAC"
mystock       <- "mac-nea"
myspeciesname <- "Mackerel"
myfile        <- file.path(datapath, "_catch_by_rectangle/New-WGWIDE catchesbyrect MAC.xlsx") 
icescatchfile <- file.path(datapath, "mac.27.nea/ICES MAC catch by country.xlsx")
acdrfile      <- "C:/TEMP/MAC compare datasets2.xlsx"

# ==========================================================================================================

# read raw data file and do conversions
catch_by_species_year_country_raw <-
  read_excel(myfile, col_names=TRUE, col_types="text") %>% 
  lowcase() %>% 
  # dplyr::mutate(across(c("catch", "lat","lon"), as.numeric ))  %>%
  dplyr::mutate(across(c("catch"), as.numeric ))  %>%
  dplyr::mutate(across(c("year"), as.integer)) %>% 
  dplyr::mutate(pnum = as.integer(gsub("quarter","", pnum))) %>% 
  dplyr::mutate(ptype = toupper(ptype)) %>% 
        
  filter(catch >0) %>% 
  
# replace lat long with values from icesrect (bottom-left points of rectangles)
  dplyr::select(-lat, -lon) %>% 
  dplyr::left_join(icesrect, by=c("rect")) %>% 
  dplyr::mutate(eez      = geo_inside(lon=lon, lat=lat, map=eez, variable="Territory1")) %>%
  
  # replace country names
  dplyr::mutate(country = ifelse(country %in% c("UKE","GBR.E") , "GBR.EW", country)) %>% 
  dplyr::mutate(country = ifelse(country %in% c("IMN","GUY","JEY","GGY"),paste("GBR",country,sep="."),country)) %>% 
  dplyr::mutate(country = ifelse(country %in% c("BES","SPA")      , "ESP", country)) %>% 
  dplyr::mutate(country = ifelse(country %in% c("POR")            , "PRT", country)) %>% 
  dplyr::mutate(country = ifelse(country %in% c("GER")            , "DEU", country)) %>% 
  
  dplyr::mutate(country = ifelse(grepl("GBR", country), "UK", country))  %>% 
  
  dplyr::mutate(
    coastalstate = case_when(
      country %in% c("IRL","ESP","NLD","DNK","DEU","FRA","PRT","SWE","POL","LTU","EST","BEL") ~ "EU27",
      TRUE ~ country
  ) )

# catch_by_species_year_country_raw %>% distinct(coastalstate) %>% View()

catch_by_species_year_country_raw_zbycountry <-
  catch_by_species_year_country_raw %>% 
  dplyr::select(-eez) %>% 
  left_join(dplyr::select(zbycountry,
                          rect, zone, area, prop), 
            by="rect") %>% 
  mutate(catch = catch * prop)

# catch_by_species_year_country_raw_zbycountry %>% distinct(zone) %>% View()

catch_by_species_year_country_raw_zbycoastalstate <-
  catch_by_species_year_country_raw %>% 
  dplyr::select(-eez) %>% 
  left_join(dplyr::select(zbycoastalstate,
                          rect, zone, area, prop), 
            by="rect") %>% 
  mutate(catch = catch * prop)

# catch_by_species_year_country_raw_zbycountry %>% distinct(country) %>% View()

# catch by year
catch_by_species_year <- 
  catch_by_species_year_country_raw %>% 
  group_by(species, year, rect, lat, lon) %>% 
  summarize(catch = sum(catch, na.rm=TRUE)) %>% 
  
  mutate(catch_interval = cut(as.integer(catch), breaks=c(1 %o% 10^(0:12)), dig.lab=10 ) ) %>% 
  mutate(catch_interval = gsub(" ","", catch_interval)) %>% 
  filter(!is.na(catch_interval)) %>% 
  filter(grepl("\\,", catch_interval)) %>%   
  group_by(species, year, rect, lat, lon, catch_interval) %>% 
  summarize(catch = sum(catch, na.rm=TRUE))

   # map_data("worldHires", xlim = xlim, ylim = ylim) %>%

compare <- 
  iAdvice %>% 
  filter(stockkeylabelold==mystock) %>% 
  filter(adviceonstock==TRUE) %>% 
  dplyr::select(year=tacyear, advice=advisedlandingsmax, 
                officiallandings, landings_ices=landings, catch_ices=catches) %>% 
  left_join(catch_by_species_year %>% 
              group_by(year) %>% 
              summarise(catch_byrect=sum(catch, na.rm=TRUE)), 
            by="year") %>% 
  
  drop_na(catch_byrect) %>% 
  
  mutate(prop_catch = catch_byrect/catch_ices) %>% 
  mutate(prop_catch2 = scales::percent(prop_catch)) %>% 
  
  mutate(prop_landings = catch_byrect/landings_ices) %>% 
  mutate(prop_landings2 = scales::percent(prop_landings))

# Catch by country dataseries (Mostly Europe, UK)
ices <-  
  read_excel(icescatchfile, col_names=TRUE, col_types="text", sheet="Sheet2") %>%
  lowcase() %>% 
  # mutate(proportionlandings = as.numeric(proportionlandings)) %>% 
  # mutate(subtotal           = as.numeric(subtotal)) %>% 
  rename(year = country) %>% 
  mutate(year = as.numeric(year))

ices2 <-
  ices %>% 
  dplyr::select(-one_of("misreported", "unallocated", "discards", "proportionlandings", "subtotal", "total")) %>% 
  tidyr::pivot_longer(names_to = "country", values_to = "catch", 2:ncol(.)) %>% 
  drop_na(catch) %>% 
  mutate(
    catch = as.numeric(catch),
    country = ifelse(grepl("faroe", country), "FRO", country),
    country = ifelse(grepl("^uk|northern|jersey|isle of|guernsey|england|united", country), "UK", country),
    country = ifelse(grepl("spain", country), "ESP", country),
    country = ifelse(grepl("ussr|russia", country), "RUS", country),
    country = ifelse(grepl("germany", country), "DEU", country), 
    country = ifelse(grepl("neth", country), "NLD", country), 
    country = ifelse(grepl("portu", country), "PRT", country), 
    country = ifelse(grepl("lith", country), "LTU", country), 
    country = ifelse(grepl("belgium", country), "BEL", country), 
    country = ifelse(grepl("france", country), "FRA", country), 
    country = ifelse(grepl("denmark", country), "DNK", country), 
    country = ifelse(grepl("sweden", country), "SWE", country), 
    country = ifelse(grepl("ireland", country), "IRL", country), 
    country = ifelse(grepl("norway", country), "NOR", country), 
    country = ifelse(grepl("eston", country), "EST", country), 
    country = ifelse(grepl("iceland", country), "ISL", country), 
    country = ifelse(grepl("greenland", country), "GRL", country), 
    country = ifelse(grepl("latvia", country), "LVA", country), 
    country = ifelse(grepl("romania", country), "ROM", country), 
    country = ifelse(grepl("poland", country), "POL", country), 
   
  ) %>% 
  group_by(year, country) %>% 
  summarise(catch = sum(catch, na.rm=TRUE)) %>% 
  mutate(dataset="ICES2") %>% 
  mutate(coastalstate = case_when(
    country == "UK" ~ "UK",
    country %in% c("BEL","DEU","DNK","NLD","FRA","POL","LTU","LVA","ESP","PRT","SWE","IRL","EST","ROM") ~ "EU27",
    TRUE ~ country
  ))

acdr <-
  read_excel(acdrfile, col_names=TRUE, col_types="text", sheet="Sheet2") %>%
  lowcase() %>%
  mutate(catch = as.numeric(catch)) %>%
  mutate(year  = as.integer(year)) %>%
  mutate(coastalstate = ifelse(country %in% c("GBR", "UK"),"UK","EU27")) %>%
  filter(dataset == "ACDR")

compare2 <-
  acdr %>%
  bind_rows(ices2) %>% 
  bind_rows(
    catch_by_species_year_country_raw %>% 
      group_by(coastalstate, country, year) %>% 
      summarise(catch = sum(catch, na.rm=TRUE)) %>% 
      mutate(dataset = "ICESRECT")
  )

converter <-
  data.frame(
    zone  = c("EU27","FRO","FRO_ISL","GBR","GBR_FRO","GRL","INT","ISL","NOR","NOR_EU27","RUS","SJM","SVA"),
    zone2 = c("EU27","FO","FO_IC","UK","UK_FO","GR","INT","IC","NO","NO_EU27","RU","SJM","SVA"),
    stringsAsFactors = FALSE)


# read the data file from Jan for checking
summarized_file <-
  read_excel("C:/DATA/Onedrive - PFA/Documents/iWGWIDE/2021/06. Data/_catch_by_rectangle/New-WGWIDE_CbyRbyMbyZone_MAC2006-2020.xlsx", 
             col_names=TRUE, 
             col_types="text") %>% 
  lowcase() %>%
  mutate(across(c(catch), as.numeric)) %>% 
  mutate(across(c(year, month), as.integer)) %>% 
  mutate(filename = "CbyRbyMbyZone")
  

# ================================================================================
# FUNCTIONS
# ================================================================================

table_catch_by_year_country <- function(myspecies="MAC", myyears=NA) {
  
  # t <-
    catch_by_species_year_country_raw %>% 
    filter(species == myspecies) %>% 
    {if ((length(myyears) == 1) & all(is.na(myyears))) {
      filter(.)
    } else {
      filter(., year %in% myyears)
    }} %>% 

    group_by(species, year, country) %>% 
    summarize(catch = as.integer(sum(catch, na.rm=TRUE))) %>% 
    reshape2::dcast(country ~ year, value.var="catch", sum, margins=c("year","country")) %>% 
    # { if (ncol(.) > split) }
    pander::pandoc.table(.,
               style = "simple",
               split.tables=120, 
               justify = "right",
               missing=".")

} # end of table_catch_by_year_country

# table_catch_by_year_country (myspecies=myspecies)

table_catch_by_year_species <- function(myyears=NA) {
  
  catch_by_species_year_country_raw %>% 
    {if ((length(myyears) == 1) & all(is.na(myyears))) {
      filter(.)
    } else {
      filter(., year %in% myyears)
    }} %>% 

    group_by(species, year) %>% 
    summarize(catch = as.integer(sum(catch, na.rm=TRUE))) %>% 
    reshape2::dcast(species ~ year, value.var="catch", sum, margins=c("year","country")) %>% 
    # { if (ncol(.) > split) }
    pander::pandoc.table(.,
               style = "simple",
               split.tables=120, 
               justify = "right",
               missing=".")

} # end of table_catch_by_year_species

# table_catch_by_year_species()


# myspecies = "HOM"; myyears = 1998:2020; ncol=6
# myspecies = "MAC"; myyears = 1998
# myspecies = "MAC"; myyears = NA
# myspecies = "HOM"; myyears = NA

plot_catch_by_year <- function(myspecies="MAC", myyears=NA, plot_catch=TRUE, plot_ssb=FALSE, ncol=6, xlim=NA, ylim=NA) {
  
  catch2 <-
    catch_by_species_year %>% 
    filter(species == myspecies) %>% 
    {if ((length(myyears) == 1) & all(is.na(myyears))) {
      filter(.)
    } else {
      filter(., year %in% myyears)
    }} %>% 
    
    {if (!all(is.na(myyears))) bind_rows(., data.frame(species=myspecies, year = myyears, catch_interval="(1,10]", catch=0)) else (.) }

  if(all(is.na(xlim))) {xlim <- range(catch2$lon, na.rm=TRUE)}
  if(all(is.na(ylim))) {ylim <- range(catch2$lat, na.rm=TRUE)}

  # catch2 %>% 
  #   group_by(year) %>% 
  #   summarise(
  #     catch=sum(catch,na.rm=TRUE), 
  #     latmin = min(lat, na.rm=TRUE), 
  #     latmax = max(lat, na.rm=TRUE), 
  #     lonmin = min(lon, na.rm=TRUE),
  #     lonmax = max(lon, na.rm=TRUE)
  #   ) %>% 
  #   View()
  
  # catch2 %>% filter(year==2018) %>% View()
  # catch_by_species_year_country %>%  filter(species=="HOM", lon == -41) %>% View()
  # hist(catch_by_species_year_country_raw$lat)
  
  tc <-
    catch2 %>% 
    group_by(species, year) %>% 
    summarize(catch = sum(catch, na.rm=TRUE)) %>% 
    mutate(catch  = as.integer(catch) ) %>% 
    
    group_by(species) %>% 
    mutate(catch2 = ylim[1] + (catch / max(catch, na.rm=TRUE) * (ylim[2] - ylim[1])) )
  
  ggplot2::map_data("world", xlim = xlim, ylim = ylim) %>% 
    ggplot(aes(long, lat)) +
    geom_polygon(aes(long, lat, group = group), fill = "gray") +
    coord_quickmap(xlim = xlim, ylim = ylim, expand = FALSE) +
    
    theme_publication() +
    theme(panel.border     = element_rect(colour="black" , size=0.2),
          panel.grid.major = element_blank(),
          strip.background = element_rect(colour="black", size =0.2),
          plot.margin      = unit(c(0,0,0,0),"cm"),
          plot.title       = element_text(hjust=0, vjust=0, size=10),
          axis.text        = element_text(size=6),
          legend.key.width = unit(0.4, "cm"), 
          axis.title       = element_blank()) +
    
    geom_tile(data=catch2, aes(x=lon, y=lat, fill = catch_interval, height=0.5, width=1), colour=NA, alpha=1.0) +
    scale_fill_viridis(option = "plasma", direction = -1, discrete=TRUE) +
    
    geom_label(data=tc, aes(label=paste0("C ",catch)), x=-Inf, y=Inf, hjust=0, vjust=1, inherit.aes=FALSE) +
  
    labs(x = NULL, y = NULL, size = "tons", title=unique(catch2$species)) +
    guides(fill = guide_legend(nrow = 1, title="catch (tonnes)")) +
    facet_wrap( ~ year, ncol=ncol)

} # end of plot_catch_by_year


plot_catch_by_year_and_country <- function() {
  
} # end of plot_catch_by_year_and_country

plot_catch_by_year_animated <- function() {
  
} # end of plot_catch_by_year_animated

# eez.df %>% 
#   filter(long >= -40, long <= 20, lat >= 35, lat <= 80) %>% 
#   filter(grepl("^IN",ISO_Ter1)) %>% 
#   distinct(Territory1, ISO_Ter1) %>% 
#   arrange(ISO_Ter1) %>% 
#   pandoc.table(style="simple", justify="left")


```

Working document xx, Coastal States Working Group, February 2022

**Full time-series of `r myspeciesname` catch by rectangle**

Martin Pastoors, `r format(Sys.time(), '%d/%m/%Y')`

**Introduction**

WGWIDE and its precursors have been publishing catch per rectangle plots in their reports for many years already. Catch by rectangle has been compiled by WG members and generally provide a WG estimate of catch per rectangle. In most cases the information is available by month or by quarter. This document provides an overview of the available meta-data from the catch by rectangle database for `r myspeciesname`.  

**Results**

**Catches of `r myspeciesname` by Coastal State and year from catch by rectangle data**

Catches of `r myspeciesname` by Coastal State and by year. Catch by rectangle data for `r myspeciesname` is only available by quarter. For 2012, only data for all countries combined is available. 

```{r, echo=FALSE, fig.asp=1.4, fig.align="center", message=FALSE, warning=FALSE}

t <-
  catch_by_species_year_country_raw %>% 
  filter(species == myspecies) %>% 
  group_by(species, year, ptype, coastalstate) %>% 
  summarize(catch = as.integer(sum(catch, na.rm=TRUE)))
  
  # group_by(species, country) %>% 
  # summarize(catch = sum(catch, na.rm=TRUE)) %>% 
  # arrange(desc(catch)) %>% 
  # View()

if(nrow(filter(t, ptype == "Q")) > 0) {
  
  cat("By quarter")
  
  t %>% 
    filter(ptype=="Q") %>% 
    group_by(species, year, coastalstate) %>% 
    summarize(catch = as.integer(sum(catch, na.rm=TRUE))) %>% 
    reshape2::dcast(year ~ coastalstate, value.var="catch", sum, margins=c("coastalstate")) %>% 
    # { if (ncol(.) > split) }
    pander::pandoc.table(.,
               style = "simple",
               split.tables=200, 
               justify = "right",
               missing=".")
}

if(nrow(filter(t, ptype == "M")) > 0) {
  
  cat("By month")
  
  t %>%
    filter(ptype=="M") %>%
    group_by(species, year, coastalstate) %>%
    summarize(catch = as.integer(sum(catch, na.rm=TRUE))) %>%
    reshape2::dcast(year ~ coastalstate, value.var="catch", sum, margins=c("coastalstate")) %>%
    # { if (ncol(.) > split) }
    pander::pandoc.table(.,
               style = "simple",
               split.tables=200,
               justify = "right",
               missing=".")
}





```

_Table 1: summed catch of `r myspeciesname` by Coastal States and year, separated into information by month (_M) or quarter (_Q)._

\newpage

**Catches of `r myspeciesname` by Country (within EU27) and year from catch by rectangle data**


```{r, echo=FALSE, fig.asp=1.4, fig.align="center", message=FALSE, warning=FALSE}

t <-
  catch_by_species_year_country_raw %>% 
  filter(coastalstate == "EU27") %>% 
  filter(species == myspecies) %>% 
  group_by(species, year, ptype, country) %>% 
  summarize(catch = as.integer(sum(catch, na.rm=TRUE)))

  
if(nrow(filter(t, ptype == "Q")) > 0) {
  
  cat("By quarter")
  
  t %>% 
    filter(ptype == "Q") %>% 
    reshape2::dcast(year ~ country, value.var="catch", sum, margins=c("country")) %>% 
    pander::pandoc.table(.,
               style = "simple",
               split.tables=200, 
               justify = "right",
               missing=".")
}

if(nrow(filter(t, ptype == "M")) > 0) {
  
  cat("By month")
  
  t %>%
    filter(ptype == "M") %>%
    reshape2::dcast(year ~ country, value.var="catch", sum, margins=c("country")) %>%
    pander::pandoc.table(.,
               style = "simple",
               split.tables=200,
               justify = "right",
               missing=".")
}




```

_Table 2: summed catch of `r myspeciesname` by EU country and year, separated into information by month (_M) or quarter (_Q)._

\newpage

**Catches of `r myspeciesname` by zone and year from catch by rectangle data 2006-2020**

```{r, echo=FALSE, fig.asp=1.4, fig.align="center", message=FALSE, warning=FALSE}

# Table 5.3.1 and figure 5.3.1 in report

panderOptions('round', 2)
panderOptions('keep.trailing.zeros', TRUE)

converter <-
  data.frame(
    zone  = c("EU27","FRO","FRO_ISL","GBR","GBR_FRO","GRL","INT","ISL","NOR","NOR_EU27","RUS","SJM","SVA"),
    zone2 = c("EU27","FO","FO_IC","UK","UK_FO","GR","INT","IC","NO","NO_EU27","RU","SJM","SVA"),
    stringsAsFactors = FALSE)

t <-
  catch_by_species_year_country_raw_zbycoastalstate %>% 
  
  filter(year >= 2006) %>% 
  filter(ptype=="M") %>% 
  drop_na(zone) %>% 
  
  left_join(converter, by="zone") %>% 
  dplyr::select(-zone) %>% 
  rename(zone=zone2) %>% 
  filter(species == myspecies) %>% 
  group_by(species, year, zone) %>% 
  summarize(catch = as.integer(sum(catch, na.rm=TRUE))) %>% 
  group_by(species, year) %>% 
  mutate(prop = catch / sum(catch, na.rm=TRUE)) 

# t %>% filter(year>= 2006) %>% group_by() %>% summarise(catch=sum(catch,na.rm=TRUE))

# table of catches by year and zone
t %>% 
  filter(zone %notin% c("RU","NO_EU27")) %>%
  reshape2::dcast(year ~ zone, value.var="catch", sum, margins=c("zone")) %>% 
  pander::pandoc.table(.,
             style = "simple",
             split.tables=200, 
             justify = "right",
             missing=".")

# write table of catches by year and zone to excel
t %>% 
  filter(zone %notin% c("RU","NO_EU27")) %>%
  reshape2::dcast(year ~ zone, value.var="catch", sum, margins=c("zone")) %>% 
  writexl::write_xlsx(., path="catch by zone and year.xlsx")
  # pander::pandoc.table(.,
  #            style = "simple",
  #            split.tables=200, 
  #            justify = "right",
  #            missing=".")

# table of prop of catches by year and zone
t %>% 
  filter(zone %notin% c("RU","NO_EU27")) %>% 
  mutate(prop = scales::percent(prop, accuracy=0.1)) %>%
  dplyr::select(-catch) %>% 
  pivot_wider(names_from = zone, values_from = prop) %>% 
  pander::pandoc.table(.,
             style = "simple",
             split.tables=200, 
             justify = "right",
             missing=".")

# table of prop of catches by year and zone to excel
t %>% 
  filter(zone %notin% c("RU","NO_EU27")) %>%
  dplyr::select(-catch) %>% 
  pivot_wider(names_from = zone, values_from = prop) %>% 
  writexl::write_xlsx(., path="catch prop by zone and year.xlsx")
  # pander::pandoc.table(.,
  #            style = "simple",
  #            split.tables=200, 
  #            justify = "right",
  #            missing=".")

# long format table of catch and prop of catches by year and zone to excel
t %>% 
  filter(zone %notin% c("RU","NO_EU27")) %>%
  writexl::write_xlsx(., path="catch and prop by zone and year long format.xlsx")

# set last years for each zone
t2 <-
  t %>%
  filter(zone %notin% c("RU","NO_EU27")) %>%
  group_by(zone) %>% 
  filter(year == max(year, na.rm=TRUE))

  
# plot
t %>% 
  filter(zone %notin% c("RU","NO_EU27")) %>%
  ggplot(aes(x=year,y=prop)) +
  theme_bw() +
  theme(legend.position="right") +
  geom_line(aes(colour=zone), size=1.0) +
  geom_point(aes(colour=zone)) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 5L)) +
  scale_x_continuous(breaks=seq(min(t$year),max(t$year)+2,2), expand  = expansion(add = c(1, 2))) +
  ggrepel::geom_text_repel(
    data=t2, 
    aes(label=zone, colour=zone), 
    # aes(label=paste(zone, scales::percent(prop, accuracy = 0.1)), colour=zone), 
    nudge_x = 0.5, direction="y", hjust="left", show.legend=FALSE) +
  # scale_size_manual(values=c("normal"=0.5,"highlight"=1.0)) +
  labs(y="relative catch by zone (%)") +
  expand_limits(y=0) +
  guides(size = "none")  
  # facet_wrap(~zone)


```

\newpage

**Catches of `r myspeciesname` by zone, country and year from catch by rectangle data**

```{r, echo=FALSE, fig.asp=1.4, fig.align="center", message=FALSE, warning=FALSE}

converter <-
  data.frame(
    zone  = c("EU27","FRO","FRO_ISL","GBR","GBR_FRO","GRL","INT","ISL","NOR","NOR_EU27","RUS","SJM","SVA"),
    zone2 = c("EU27","FO","FO_IC","UK","UK_FO","GR","INT","IC","NO","NO_EU27","RU","SJM","SVA"),
    stringsAsFactors = FALSE)


t <-
  catch_by_species_year_country_raw_zbycoastalstate %>% 
  
  filter(year >= 2006) %>% 
  filter(ptype=="M") %>%
  drop_na(zone) %>% 
  
  left_join(converter, by="zone") %>% 
  dplyr::select(-zone) %>% 
  rename(zone=zone2, month=pnum) %>% 
  filter(species == myspecies) %>% 
  
  # first sum by month to make the tables consistent
  group_by(species, year, month, zone) %>% 
  summarize(catch = (sum(catch, na.rm=TRUE))) %>% 
  # summarize(catch = as.integer(sum(catch, na.rm=TRUE))) %>% 
  
  filter(catch > 0) %>% 
  
  group_by(species, year, zone) %>% 
  mutate(prop = catch / sum(catch, na.rm=TRUE)) 

# t %>% filter(year>= 2006) %>% group_by() %>% summarise(catch=sum(catch,na.rm=TRUE))

t %>% 
  filter(zone %notin% c("RU","NO_EU27")) %>%
  writexl::write_xlsx(., path="catch and prop by zone and year and month long format.xlsx")


# t %>% 
#   filter(zone %notin% c("RU","NO_EU27")) %>%
#   mutate(prop = scales::percent(prop, accuracy=0.1)) %>%
#   dplyr::select(-catch) %>% 
#   pivot_wider(names_from = zone, values_from = prop) %>% 
#   pander::pandoc.table(.,
#              style = "simple",
#              split.tables=200, 
#              justify = "right",
#              missing=".")

# set last years for each zone and month
t2 <-
  t %>%
  filter(zone %notin% c("RU","NO_EU27")) %>%
  group_by(month, zone) %>% 
  filter(year == max(year, na.rm=TRUE))

  
# plot
t %>% 
  filter(zone %notin% c("RU","NO_EU27")) %>%
  ggplot(aes(x=year,y=prop)) +
  theme_bw() +
  theme(legend.position="right") +
  geom_line(aes(colour=zone), size=1.0) +
  geom_point(aes(colour=zone)) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 5L)) +
  scale_x_continuous(breaks=seq(min(t$year),max(t$year)+2,2), expand  = expansion(add = c(1, 4))) +
  ggrepel::geom_text_repel(
    data=t2, 
    aes(label=zone, colour=zone), 
    # aes(label=paste(zone, scales::percent(prop, accuracy = 0.1)), colour=zone), 
    nudge_x = 0.5, direction="y", hjust="left", show.legend=FALSE) +
  # scale_size_manual(values=c("normal"=0.5,"highlight"=1.0)) +
  labs(y="relative catch by zone (%)") +
  expand_limits(y=0) +
  guides(size = "none")  +
  facet_wrap(~month)



```
```{r, echo=FALSE, fig.asp=1.4, fig.align="center", message=FALSE, warning=FALSE}

# same but also by country

converter <-
  data.frame(
    zone  = c("EU27","FRO","FRO_ISL","GBR","GBR_FRO","GRL","INT","ISL","NOR","NOR_EU27","RUS","SJM","SVA"),
    zone2 = c("EU27","FO","FO_IC","UK","UK_FO","GR","INT","IC","NO","NO_EU27","RU","SJM","SVA"),
    stringsAsFactors = FALSE)


t <-
  catch_by_species_year_country_raw_zbycoastalstate %>% 
  
  filter(year >= 2006) %>% 
  filter(ptype=="M") %>%
  drop_na(zone) %>% 
  
  left_join(converter, by="zone") %>% 
  dplyr::select(-zone) %>% 
  rename(zone=zone2, month=pnum) %>% 
  filter(species == myspecies) %>% 
  
  # first sum by month to make the tables consistent
  group_by(species, year, month, country, zone) %>% 
  summarize(catch = (sum(catch, na.rm=TRUE))) %>% 
  
  filter(catch > 0) %>% 
  
  group_by(species, year, country, zone) %>% 
  mutate(prop = catch / sum(catch, na.rm=TRUE)) 

# t %>% filter(year>= 2006) %>% group_by() %>% summarise(catch=sum(catch,na.rm=TRUE))

t %>% 
  filter(zone %notin% c("RU","NO_EU27")) %>%
  writexl::write_xlsx(., path="catch and prop by country and zone and year and month long format.xlsx")


# t %>% 
#   filter(zone %notin% c("RU","NO_EU27")) %>%
#   mutate(prop = scales::percent(prop, accuracy=0.1)) %>%
#   dplyr::select(-catch) %>% 
#   pivot_wider(names_from = zone, values_from = prop) %>% 
#   pander::pandoc.table(.,
#              style = "simple",
#              split.tables=200, 
#              justify = "right",
#              missing=".")

# set last years for each zone and month
t2 <-
  t %>%
  filter(zone %notin% c("RU","NO_EU27")) %>%
  group_by(month, zone) %>% 
  filter(year == max(year, na.rm=TRUE))

  
# plot
t %>% 
  filter(zone %notin% c("RU","NO_EU27")) %>%
  ggplot(aes(x=year,y=prop)) +
  theme_bw() +
  theme(legend.position="right") +
  geom_line(aes(colour=zone), size=1.0) +
  geom_point(aes(colour=zone)) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 5L)) +
  scale_x_continuous(breaks=seq(min(t$year),max(t$year)+2,2), expand  = expansion(add = c(1, 4))) +
  ggrepel::geom_text_repel(
    data=t2, 
    aes(label=zone, colour=zone), 
    # aes(label=paste(zone, scales::percent(prop, accuracy = 0.1)), colour=zone), 
    nudge_x = 0.5, direction="y", hjust="left", show.legend=FALSE) +
  # scale_size_manual(values=c("normal"=0.5,"highlight"=1.0)) +
  labs(y="relative catch by zone (%)") +
  expand_limits(y=0) +
  guides(size = "none")  +
  facet_wrap(~month)



```

\newpage

**Comparison of landings or catches of `r myspeciesname` from different data sources**

A comparison is made of summed catch by rectangle data (red), ICES catch estimates (green) and ICES landings estimates (blue).

```{r, echo=FALSE, fig.asp=0.6, fig.align="center", message=FALSE, warning=FALSE}


compare %>% 
  dplyr::select(year, landings_ices, catch_ices, catch_byrect) %>% 
  pivot_longer(names_to = "variable", values_to = "data", c(landings_ices, catch_ices, catch_byrect)) %>% 
  ggplot(aes(x=year,y=data)) +
  theme_publication() +
  geom_line(aes(colour=variable)) +
  geom_point(aes(colour=variable)) +
  labs(y="catch/landings (tonnes)") +
  expand_limits(y=0)


```

_Figure 1: Comparison of summed catch by rectangle data (red), ICES catch estimates (green) and ICES landings estimates (blue)._

\newpage


**Comparison of landings of `r myspeciesname` of EU27 countries from different data sources**


```{r, echo=FALSE, fig.asp=1.2, fig.align="center", message=FALSE, warning=FALSE}


compare2 %>% 
  filter(coastalstate == "EU27") %>% 
  filter(year >= 2012) %>% 
  
  ggplot(aes(x=year,y=catch)) +
  theme_publication() +
  geom_line(aes(colour=dataset)) +
  geom_point(aes(colour=dataset)) +
  labs(y="catch (tonnes)") +
  expand_limits(y=0) +
  facet_wrap(~country, scales="free_y")



```

_Figure 2: Comparison of summed catch by rectangle data (red), ICES catch estimates (green) and ICES landings estimates (blue)._

\newpage

\newpage




```{r eval=FALSE, fig.align="center", fig.asp=1.2, message=FALSE, warning=FALSE, include=FALSE}

# catch_by_species_year_country_raw_zbyrect %>% distinct(zone) %>% View()

sf::sf_use_s2(FALSE)

t <-
  catch_by_species_year_country_raw %>% 
  filter(year >= 2006) %>% 
  group_by(species, year, lat, lon) %>% 
  summarize(catch = as.integer(sum(catch, na.rm=TRUE))) %>% 
  mutate(catch2 = cut(catch, breaks=c(0,100,1000,5000,25000,500000),  dig.lab=10))

# unique(t$catch2)

xlim <- range(t$lon, na.rm=TRUE)
ylim <- range(t$lat, na.rm=TRUE)

ggplot2::map_data("world", xlim = xlim, ylim = ylim) %>% 
  ggplot(aes(long, lat)) +
  
  theme_publication() +
  theme(panel.border     = element_rect(colour="black" , size=0.2),
        panel.grid.major = element_blank(),
        strip.background = element_rect(colour="black", size =0.2),
        plot.margin      = unit(c(0,0,0,0),"cm"),
        plot.title       = element_text(hjust=0, vjust=0, size=10),
        axis.text        = element_text(size=6),
        legend.key.width = unit(0.4, "cm"), 
        axis.title       = element_blank()) +
  
  coord_quickmap(xlim = xlim, ylim = ylim, expand = FALSE) +
  
  geom_tile(data=t, aes(x=lon, y=lat, fill = catch2, height=0.5, width=1), colour=NA, alpha=1.0) +
  scale_fill_viridis(option = "plasma", direction = -1, discrete=TRUE) +
  
  geom_polygon(aes(long, lat, group = group), fill = "gray") +
  geom_polygon(data=eez.df, aes(long, lat, group = group), fill = NA, colour="gray") +
  
  labs(x = NULL, y = NULL, size = "tons", title="MAC catch by rectangle by year") +
  guides(fill = guide_legend(nrow = 1, title="catch (tonnes)")) +
  facet_wrap( ~ year, ncol=4)



```


```{r, echo=FALSE, fig.asp=0.6, fig.align="center", message=FALSE, warning=FALSE}

t <-
  catch_by_species_year_country_raw_zbycoastalstate %>% 
  filter(year >= 2006) %>% 
  filter(!is.na(zone)) %>% 
  filter(ptype == "M") %>% 
  rename(month=pnum) %>% 
  group_by(species, year, month) %>% 
  summarize(catch = as.integer(sum(catch, na.rm=TRUE))) %>% 
  group_by(species, year) %>%
  mutate(prop = catch / sum(catch,na.rm=TRUE)) %>% 
  mutate(period = cut(year, breaks=c(2005,2010, 2015, 2020)))
  

t %>% 
  ggplot(aes(x=month,y=prop)) +
  theme_publication() +
  theme(legend.position="right") +
  geom_line(aes(group=year)) +
  geom_smooth() +
  scale_y_continuous(labels = scales::percent_format(accuracy = 5L)) +
  scale_x_continuous(breaks=seq(1,12,1)) +
  labs(y="relative catch by month (%)") +
  expand_limits(y=0) +
  facet_wrap(~period, ncol=1)



```

```{r eval=FALSE, fig.align="center", fig.asp=1.2, message=FALSE, warning=FALSE, include=FALSE}

# proportion of catches in rectangles that are split between different zones

t <-
  catch_by_species_year_country_raw_zbycoastalstate %>% 
  filter(year >= 2006) %>% 
  group_by(species, year, rect, zone) %>% 
  summarize(catch = as.integer(sum(catch, na.rm=TRUE))) %>% 
  group_by(species, year, rect) %>% 
  mutate(n=n()) %>% 
  group_by(species, year, n) %>% 
  summarize(catch = as.integer(sum(catch, na.rm=TRUE))) %>% 
  group_by(species, year) %>%
  mutate(prop= catch / sum(catch, na.rm=TRUE)) %>% 
  mutate(n=paste(n,"zone(s)"))
  
t2 <-
  t %>% 
  group_by(species, n) %>% 
  summarise(prop = mean(prop, na.rm=TRUE)) %>% 
  mutate(perc = scales::percent(prop, accuracy = 0.1))


sum(t2$prop)

t %>% 
  ggplot(aes(x=year,y=prop)) +
  theme_bw() +
  theme(legend.position="right") +
  geom_line(aes(colour=n), size=1) +
  geom_point(aes(colour=n)) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 5L)) +
  scale_x_continuous(breaks=seq(min(t$year),max(t$year),2)) +
  labs(y="percentage catch with x zones per rect") +
  expand_limits(y=0) +
  guides(size = "none")  


```