---
output: 
  word_document:
    reference_docx: C:\Users\Martin Pastoors\AppData\Roaming\Microsoft\QuickStyles\Rmarkdown style.dotx
---

```{r setup, include=FALSE}

require("knitr")
knitr::opts_chunk$set(
	echo = FALSE,
	message = FALSE,
	warning = FALSE,
	crop = TRUE,
	comment = ""
)
knit_hooks$set(crop = hook_pdfcrop)
opts_knit$set(root.dir="D:/XXX/PRF/")
```


```{r initialization, include=FALSE}
# =======================================================================================
# Stock summaries.Rmd
# 
# Generate stock summary report for WGWIDE
#
# 03/09/2017 first version
# =======================================================================================

rm(list=ls())

# Libraries
library(rmarkdown)
library(maps)          # world maps
library(rgdal)         # read spatial polygons (e.g. EEZ)
library(sp)            # spatial data 
library(geo)           # spatial manipulation
library(pander)        # tables
library(colorspace)    # hcl colours
library(cowplot)       # combine different plots. 

library(tidyverse)     # combined package of dplyr, tidyr, ggplot, readr, purrr and tibble
library(lubridate)     # data handling
library(reshape2)      # reshaping data; e.g. cast
library(readxl)        # excel reader
library(broom)         # clean up statistics
library(scales)        # pretty scales

options(max.print=999999)

path    <- "Y:/2017 Meeting docs/06. Data/stock summaries"
setwd(path)

# Load utils code
source("D:/GIT/mptools/r/my_utils.r")

mac <- read_excel(path="stock summaries wgwide 2017.xlsx", sheet="mac", skip=2, col_names=T) %>% 
  lowcase() %>% 
  mutate(species = "mac") %>% 
  select(species, year, r, ssb, f) 

whb <- read_excel(path="stock summaries wgwide 2017.xlsx", sheet="whb", skip=2, col_names=T) %>% 
  lowcase() %>% 
  mutate(species = "whb") %>% 
  select(species, year, r, ssb, f) 

her <- read_excel(path="stock summaries wgwide 2017.xlsx", sheet="her", skip=2, col_names=T) %>% 
  lowcase() %>% 
  mutate(species = "her",
         ssb     = ssb * 1000,
         r       = r * 1000)%>% 
  select(species, year, r, ssb, f) 

hom <- read_excel(path="stock summaries wgwide 2017.xlsx", sheet="hom", skip=2, col_names=T) %>% 
  lowcase() %>% 
  mutate(species = "hom")%>% 
  select(species, year, r, ssb, f) 

summ <- rbind(mac, whb, her, hom)

# calculating averages
avg <-
  summ %>% 
  gather(key=var, value=value, r:f) %>% 
  group_by(species, var) %>% 
  summarise(mean    = mean(value, na.rm=TRUE),
            geomean = exp(mean(log(value), na.rm=TRUE)),
            median  = median(value, na.rm=TRUE)) %>% 
  mutate(mean = ifelse(var == "r", geomean, mean)) %>% 
  select(-geomean)


```

# Stock summaries WGWIDE 2017

&nbsp;  

Martin Pastoors, `r format(Sys.time(), '%d/%m/%Y')`

&nbsp;  
&nbsp;  

## Introduction

Providing an overview of pelagic stock trends in the northeast Atlantic

Combining results of assessments of Norwegian Spring Spawning Herring, Blue whiting, Northeast Atlantic mackerel and Horse mackerel. 

##### page break

## SSB

Spawning stock biomass of four stocks both on absolute scale (tonnes) and relative scale (only for the year in common). 

_Figure 1: SSB of mackerel, horse mackerel, blue whiting and herring_

```{r ssb_abs, echo=FALSE, fig.width=10, fig.asp=.8, fig.align="center", message=FALSE, warning=FALSE}

p1 <-
  summ %>% 
  ggplot(aes(year, ssb)) +
  theme_publication() +
  theme(plot.margin = unit(c(0.1, 0.1, 0.1, 0.1), "cm")) +
  # geom_line(aes(colour=species)) +
  geom_bar(aes(fill=species), colour=NA, stat="identity", position="stack")

p2 <-
  summ %>% 
  filter(year >= 1988) %>% 
  ggplot(aes(year, ssb)) +
  theme_publication() +
  theme(plot.margin = unit(c(0.1, 0.1, 0.1, 0.1), "cm")) +
  # geom_line(aes(colour=species)) +
  geom_bar(aes(fill=species), colour=NA, stat="identity", position="fill") +
  xlim(1980,2017)

plot_grid(p1 + theme(legend.position = "none", 
                     axis.title.x    = element_blank(),
                     axis.text.x     = element_blank()) , 
          p2,  
          ncol=1, align = 'v', rel_heights = c(0.5,0.6))
```

##### page break

Overview of the trends in fishing mortality, recruitment and SSB for the four major pelagic stocks.

Simple averages over the time series are shown for F and SSB. Geometric mean for recruitment. 

_Figure 2: F, Recruitment (R) and SSB of mackerel, horse mackerel, blue whiting and herring_

```{r stocksum, echo=FALSE, fig.width=10, fig.asp=.7, fig.align="center", message=FALSE, warning=FALSE}

summ %>% 
  gather(key=var, value=value, r:f) %>% 
  
  ggplot(aes(year, value)) +
  theme_publication() +
  theme(legend.position="none", 
        axis.title.y = element_blank()) +
  geom_line(aes(colour=species)) +
  geom_hline(data=avg, aes(colour=species, yintercept = mean), linetype="dashed") +
  facet_grid(var~species, scales="free_y")


```


