---
output: 
  word_document:
    reference_docx: ../PFA_report_template_v1.4.2.dotx
---

```{r setup, include=FALSE}

# =======================================================================================
# WGWIDE Stock summaries.Rmd
# 
# Generate stock summary report for WGWIDE
#
# 03/09/2017 first version
# 06/08/2018 UPdated and prepared for WGWIDE 2018
# 07/09/2018 Added link to SAG database for plotting. Currently reading from downloaded data. 
#            Needs to be made through live connection. 
# =======================================================================================

require("knitr")
knitr::opts_chunk$set(echo = FALSE,	message = FALSE,	warning = FALSE,	comment = "",	crop = TRUE )
knitr::opts_chunk$set(fig.width=10) 

options(max.print=1000000)

rm(list=ls())

# Libraries
library(rmarkdown)
library(readxl)        # excel reader
library(tidyverse)
library(cowplot)
library(directlabels)  # for printing labels at end of geom lines
library(scales)

# Load utils code
source("../../mptools/r/my_utils.r")

# Set dropbox folder
dropboxdir <- paste(get_dropbox(), "/ICES Assessment database", sep="")
wgwidedir  <- "d:/wgwide/2018 Meeting docs/06. Data"

# Load assessment data
sag <-
  get(load(file=file.path(dropboxdir, "rdata/icesSAGfull 20180902.RData"))) %>% 
  filter(stockkeylabelold %in% c("her-noss", "whb-comb","hom-west", "mac-nea"),
         assessmentyear == 2018, year <= 2018) %>% 
  group_by(stockkeylabelold) %>% 
  filter(assessmentkey == max(assessmentkey))


# load rbya data of the different stocks (not for her-noss so far)
rbya <- data.frame()
for (s in c("hom.27.2a4a5b6a7a-ce-k8","mac.27.nea","whb.27.1-91214")) {
  print(s)
  rbya <- bind_rows(rbya, 
                    read.csv(file=file.path(wgwidedir, s, "output/tidy/rbya.csv"), stringsAsFactors = F))
}

```

# Stock summaries WGWIDE 2018

&nbsp;  

Martin Pastoors, `r format(Sys.time(), '%d/%m/%Y')`

&nbsp;  
&nbsp;  

## Introduction

This working group has carried out the stock assessments of the following widely distributed
and migratory pelagic species: boarfish, red gurnard, Norwegian spring spawning
herring, Western horse mackerel, North Sea horse mackerel, Northeast Atlantic
mackerel, Striped red mullet and Blue whiting. 

Analytical (category 1) type of assessments are available for the four main species that make up the bulk of the biomass of pelagic species in the Northeast Atlantic:

* Northeast Atlantic mackerel
* Norwegian spring spawning herring
* Blue whiting
* Western horse mackerel.

The fluctuations in the catches of the four stocks since 1988 are shown in Figure 1.10.1.

```{r, echo=FALSE, fig.width=10, fig.asp=0.8, fig.align="center", message=FALSE, warning=FALSE}

p1 <-
  sag %>% 
  filter(year >= 1988) %>% 
  mutate(catches = ifelse(year == max(year), NA, catches)) %>% 
  
  ggplot(aes(year, catches)) +
  theme_publication() +
  theme(plot.margin = unit(c(0.1, 0.1, 0.1, 0.1), "cm")) +
  # geom_line(aes(colour=species)) +
  geom_bar(aes(fill=stockkeylabel), colour=NA, stat="identity", position="stack") +
  labs(y="catch (tonnes)") +
  scale_y_continuous(labels = comma_format())

p2 <-
  sag %>% 
  filter(year >= 1988) %>% 
  mutate(catches = ifelse(year == max(year), NA, catches)) %>% 
  
  ggplot(aes(year, catches)) +
  theme_publication() +
  theme(plot.margin = unit(c(0.1, 0.1, 0.1, 0.1), "cm")) +
  # geom_line(aes(colour=species)) +
  geom_bar(aes(fill=stockkeylabel), colour=NA, stat="identity", position="fill") +
  labs(y="rel. stocksize") +
  scale_y_continuous(labels = percent_format())

plot_grid(p1 + theme(legend.position = "none", 
                     axis.title.x    = element_blank(),
                     axis.text.x     = element_blank()) , 
          p2,  
          ncol=1, align = 'v', rel_heights = c(0.5,0.6))
```

_Figure 1.10.1: Catch of mackerel, western horse mackerel, blue whiting and Norwegian spring spawning herring_


The trends in SSB of the four stocks are shown in Figure 1.10.2, both in absolute biomass (tonnes) and in relative proportions. At the maximum, pelagic biomass of these species has been around 15 million tonnes. Recently the biomass appears to have decreased to around 12 million tonnes. The contributions of Norwegian Spring-spawning herring, Western horse mackerel and Northeast Atlantic mackerel has decreased in recent year while blue whiting has increased. 

```{r ssb_abs, echo=FALSE, fig.width=10, fig.asp=1.4, fig.align="center", message=FALSE, warning=FALSE}


p0 <-
  sag %>% 
  filter(year >= 1988) %>% 
  ggplot(aes(year, stocksize)) +
  theme_publication() +
  theme(plot.margin = unit(c(0.1, 0.1, 0.1, 0.1), "cm")) +
  geom_line(aes(colour=stockkeylabel)) +
  # geom_bar(aes(fill=stockkeylabel), colour=NA, stat="identity", position="stack") +
  labs(y="ssb (tonnes)") +
  expand_limits(y=0) +
  scale_y_continuous(labels = comma_format())

p1 <-
  sag %>% 
  filter(year >= 1988) %>% 
  ggplot(aes(year, stocksize)) +
  theme_publication() +
  theme(plot.margin = unit(c(0.1, 0.1, 0.1, 0.1), "cm")) +
  # geom_line(aes(colour=species)) +
  geom_bar(aes(fill=stockkeylabel), colour=NA, stat="identity", position="stack") +
  labs(y="ssb (tonnes)") +
  scale_y_continuous(labels = comma_format())

p2 <-
  sag %>% 
  filter(year >= 1988) %>% 
  ggplot(aes(year, stocksize)) +
  theme_publication() +
  theme(plot.margin = unit(c(0.1, 0.1, 0.1, 0.1), "cm")) +
  # geom_line(aes(colour=species)) +
  geom_bar(aes(fill=stockkeylabel), colour=NA, stat="identity", position="fill") +
  labs(y="rel. stocksize") +
  scale_y_continuous(labels = percent_format())

plot_grid(p0 + theme(legend.position = "none", 
                     axis.title.x    = element_blank(),
                     axis.text.x     = element_blank()) , 
          p1 + theme(legend.position = "none", 
                     axis.title.x    = element_blank(),
                     axis.text.x     = element_blank()) , 
          p2,  
          ncol=1, align = 'v', rel_heights = c(0.5, 0.5,0.6))
```

_Figure 1.10.2: SSB of mackerel, western horse mackerel, blue whiting and Norwegian spring spawning herring_

An overview of the key variables for each of the stocks (Fishing pressure (F), recruitment (R) and Spawning Stock Biomass (SSB)) is shown in Figure 1.10.3. From these comparisons it can be concluded that the fishing mortality of mackerel and blue whiting has generally been higher than the fishing mortality of horse mackerel and herring. Recruitment levels of blue whiting and herring are on a comparable scale and substantially higher and horse mackerel (except for the 1982 year class) and mackerel. Biomass trends of the different stocks are somewhat on the same level but show very different tendencies.

```{r, echo=FALSE, fig.width=10, fig.asp=.7, fig.align="center", message=FALSE, warning=FALSE}

avg <-
  sag %>% 
  gather(key=var, value=value, c(stocksize, fishingpressure, recruitment)) %>% 
  group_by(stockkeylabel, var) %>% 
  mutate(value = ifelse(var=="recruitment", log(value), value)) %>% 
  summarize(value = mean(value, na.rm=TRUE)) %>% 
  mutate(value = ifelse(var=="recruitment", exp(value), value)) 


sag %>% 
  gather(key=var, value=value, c(stocksize, fishingpressure, recruitment)) %>% 
  
  ggplot(aes(year, value)) +
  theme_publication() +
  theme(legend.position="none", 
        axis.title.y = element_blank()) +
  geom_line(aes(colour=stockkeylabel)) +
  geom_hline(data=avg, aes(colour=stockkeylabel, yintercept = value), linetype="dashed") +
  facet_grid(var~stockkeylabel, scales="free_y")


```

_Figure 1.10.3: SSB of mackerel, western horse mackerel, blue whiting and Norwegian spring spawning herring_

An overview of stock weight at age for mackerel and blue whiting is shown in figure 1.10.4 (NOTE: horse mackerel and Atlanto-scandian herring still missing). Older mackerel has experienced a substantially lower weight at in the recent years although this tendency appears to have changed for the younger ages. Weight at age of blue whiting shows substantial fluctuations which appear to be somewhat related to the stock size. 

```{r, echo=FALSE, fig.width=10, fig.asp=1.0, fig.align="center", message=FALSE, warning=FALSE}

# plot of stock weight at age

for (s in c("mac","whb")) {
  
  print(
  rbya %>% 
    filter(age >= 2, age <= 10) %>% 
    mutate(species = substr(stock,1,3)) %>% 
    filter(species == s) %>%   
    
    ggplot(aes(x=year, y=sW, group=species)) +
    theme_publication() +
    theme(legend.position="none")+
    geom_point(aes(colour=factor(species))) +
    geom_smooth(aes(colour=factor(species)), method="loess", span=0.2, se=FALSE) +
    facet_wrap(~age, scales="free") +
    labs(title = paste(toupper(s), "stock weight at age", sep=" "))
  )
  
}


```

